### 2020-04-06 production/prod_analysis-long_invariance.rmd
update-0604-2020
* Actualización de análisis, con foco en ola 1.
* El documento ahora está completamente en inglés.

* Análisis longitudinal se movió a un archivo independiente (prod_analysis-long-invariance.rmd

---
title: "Longitudinal analysis "
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: "../input/bib/style.css"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: "hide"
editor_options: 
  chunk_output_type: console
---
# Longitudinal analysis

```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
load(file = "input/data/proc/datawide_123.RData")
```

```{r}
g01<- wide02 %>% filter(grupo==1) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by perception/preference
g02<- wide02 %>% filter(grupo==2) %>% select(starts_with("perc_"),starts_with("pref_")) # Fixed by topic (i.e: effort)
g03<- wide02 %>% filter(grupo==3) %>% select(starts_with("perc_"),starts_with("pref_")) # Randomized
```

### Grupo Versión 01

```{r, out.width="75%", results='asis'}
stargazer(g01,type = "html", median = T,digits = 2)
```

```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
corwide01a<- polychoric(x = select(g01 ,starts_with(match = "perc_")),na.rm = T)
rownames(corwide01a$rho) <-c("1. pc_effortw1","2. pc_talentw1","3. pc_wpartw1 ","4. pc_netww1","5. pc_effortw2","6. pc_talentw2","7. pc_wpartw2","8. pc_netww2","9. pc_effortw3","10. pc_talentw3","11. pc_wpartw3 ","12. pc_netww3")
colnames(corwide01a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide01a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)

corwide01b<- polychoric(x = select(g01 ,starts_with(match = "pref_")),na.rm = T)
rownames(corwide01b$rho) <-c("1. pf_effortw1","2. pf_talentw1","3. pf_wpartw1 ","4. pf_netww1","5. pf_effortw2","6. pf_talentw2","7. pf_wpartw2","8. pf_netww2","9. pf_effortw3","10. pf_talentw3","11. pf_wpartw3 ","12. pf_netww3")
colnames(corwide01b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide01b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)

```

### Grupo Versión 02

```{r, out.width="75%", results='asis'}
stargazer(g02,type = "html", median = T,digits = 2)
```

```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
corwide02a<- polychoric(x = select(g02 ,starts_with(match = "perc_")),na.rm = T)
rownames(corwide02a$rho) <-c("1. pc_effortw1","2. pc_talentw1","3. pc_wpartw1 ","4. pc_netww1","5. pc_effortw2","6. pc_talentw2","7. pc_wpartw2","8. pc_netww2","9. pc_effortw3","10. pc_talentw3","11. pc_wpartw3 ","12. pc_netww3")
colnames(corwide02a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide02a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)

corwide02b<- polychoric(x = select(g02 ,starts_with(match = "pref_")),na.rm = T)
rownames(corwide02b$rho) <-c("1. pf_effortw1","2. pf_talentw1","3. pf_wpartw1 ","4. pf_netww1","5. pf_effortw2","6. pf_talentw2","7. pf_wpartw2","8. pf_netww2","9. pf_effortw3","10. pf_talentw3","11. pf_wpartw3 ","12. pf_netww3")
colnames(corwide02b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide02b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
```

### Grupo Versión 03

```{r, results='asis'}
stargazer(g03,type = "html", median = T,digits = 2)
```

```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
corwide03a<- polychoric(x = select(g03 ,starts_with(match = "perc_")),na.rm = T)
rownames(corwide03a$rho) <-c("1. pc_effortw1","2. pc_talentw1","3. pc_wpartw1 ","4. pc_netww1","5. pc_effortw2","6. pc_talentw2","7. pc_wpartw2","8. pc_netww2","9. pc_effortw3","10. pc_talentw3","11. pc_wpartw3 ","12. pc_netww3")
colnames(corwide03a$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide03a$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)

corwide03b<- polychoric(x = select(g03 ,starts_with(match = "pref_")),na.rm = T)
rownames(corwide03b$rho) <-c("1. pf_effortw1","2. pf_talentw1","3. pf_wpartw1 ","4. pf_netww1","5. pf_effortw2","6. pf_talentw2","7. pf_wpartw2","8. pf_netww2","9. pf_effortw3","10. pf_talentw3","11. pf_wpartw3 ","12. pf_netww3")
colnames(corwide03b$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwide03b$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
```

### Muestra completa

```{r, results='asis'}
stargazer(select(wide02 ,starts_with(match = "perc_"), starts_with("pref_")),type = "html", median = T,digits = 2)
```

```{r, out.width=c("75%","75%"), fig.cap=c("Percepción","Preferencia"), fig.align='center'}
corwideA<- polychoric(x = select(wide02 ,starts_with(match = "perc_")),na.rm = T)
rownames(corwideA$rho) <-c("1. pc_effortw1","2. pc_talentw1","3. pc_wpartw1 ","4. pc_netww1","5. pc_effortw2","6. pc_talentw2","7. pc_wpartw2","8. pc_netww2","9. pc_effortw3","10. pc_talentw3","11. pc_wpartw3 ","12. pc_netww3")
colnames(corwideA$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwideA$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)

corwideB<- polychoric(x = select(wide02 ,starts_with(match = "pref_")),na.rm = T)
rownames(corwideB$rho) <-c("1. pf_effortw1","2. pf_talentw1","3. pf_wpartw1 ","4. pf_netww1","5. pf_effortw2","6. pf_talentw2","7. pf_wpartw2","8. pf_netww2","9. pf_effortw3","10. pf_talentw3","11. pf_wpartw3 ","12. pf_netww3")
colnames(corwideB$rho) <-c("(1)", "(2)","(3)","(4)","(5)","(6)","(7)", "(8)","(9)","(10)","(11)","(12)")
corrplot(corwideB$rho,method = "color", type = "upper", tl.col = "black", addCoef.col = "black",diag = FALSE)
```

## CFA wave01, wave02 y wave03

```{r}
model_long <- '
# wave 01
perc_merit1 =~perc_effort_w1+perc_talent_w1 
perc_nmerit1=~perc_wpart_w1 +perc_netw_w1  
pref_merit1 =~pref_effort_w1+pref_talent_w1 
pref_nmerit1=~pref_wpart_w1 +pref_netw_w1
# wave 02
perc_merit2 =~perc_effort_w2+perc_talent_w2 
perc_nmerit2=~perc_wpart_w2 +perc_netw_w2  
pref_merit2 =~pref_effort_w2+pref_talent_w2 
pref_nmerit2=~pref_wpart_w2 +pref_netw_w2
# wave 03
perc_merit3 =~perc_effort_w3+perc_talent_w3 
perc_nmerit3=~perc_wpart_w3 +perc_netw_w3  
pref_merit3 =~pref_effort_w3+pref_talent_w3 
pref_nmerit3=~pref_wpart_w3 +pref_netw_w3
'

fitlong_c <- cfa(model = model_long,data = wide02,estimator="MLR") # Continuous/ estimador ML Robust

Indicatores <-c("perc_effort_w1","perc_talent_w1","perc_wpart_w1","perc_netw_w1",
                "pref_effort_w1","pref_talent_w1","pref_wpart_w1","pref_netw_w1",
                "perc_effort_w2","perc_talent_w2","perc_wpart_w2","perc_netw_w2",
                "pref_effort_w2","pref_talent_w2","pref_wpart_w2","pref_netw_w2")

fitlong_o <- cfa(model = model_long,data = wide02,ordered = Indicatores)

#---------------------------------------------------#
kable(rbind(Continuous=fitmeasures(fitlong_c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            ordinal= fitmeasures(fitlong_o)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),
      format = "markdown",digits = 3)

cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
kable(left_join(x = standardizedsolution(fitlong_c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
                y = standardizedsolution(fitlong_o) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),
                c("lhs","rhs")),
      format = "markdown",digits = 2,col.names = cnames)
```

```{r,fig.cap="Correlación entre variables latentes w01, w02 y w03"}
cormat<- as.matrix(inspect(fitlong_o, "cor.lv"))
corrplot(cormat, method = "color" ,type = "upper", tl.col = "black", tl.srt = 45,addCoef.col = "black",diag = FALSE)
```

# Invarianza

## Invarianza entre modalidades en ola 1,2 y 3 (between groups invariance)

**NOTE**: En vista de que no existe una librería que permita estimar invarianza longitudinal asumiendo Indicatores ordinales, las primeras estimaciones las realizaré asumiendo Indicatores Continuouss (estimator=MLR).

```{r}
library(semTools)
```

```{r}
load(file = "input/data/proc/datalong.RData")     # Datos formato long
load(file = "input/data/proc/datawide_123.RData") # Datos formato wide 3 olas

wide_m1 <- subset(wide02,grupo==1) # subset for mode 01
wide_m2 <- subset(wide02,grupo==2) # subset for mode 02
wide_m3 <- subset(wide02,grupo==3) # subset for mode 03
```

```{r}
m_inv_modalidad <- 'perc_merit=~ perc_effort+perc_talent 
                    perc_nmerit=~perc_wpart +perc_netw  
                    pref_merit=~ pref_effort+pref_talent 
                    pref_nmerit=~pref_wpart +pref_netw'

indicators  <-c("perc_effort","perc_talent","perc_wpart","perc_netw",
                "pref_effort","pref_talent","pref_wpart","pref_netw")

fit.conf=cfa(m_inv_modalidad,data=dat04,group="grupo",estimator="MLR")
fit.deb =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings"),estimator="MLR")
fit.fuer=cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts"),estimator="MLR")
fit.str =cfa(m_inv_modalidad,data=dat04,group="grupo",group.equal=c("loadings","intercepts","residuals"),estimator="MLR")
```

```{r}
kable(rbind(fit.conf=fitmeasures(fit.conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.deb =fitmeasures(fit.deb )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.fuer=fitmeasures(fit.fuer)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.str =fitmeasures(fit.str )[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")

kable(anova(fit.conf, fit.deb, fit.fuer, fit.str),format = "markdown",digits = 3)
```

## Invarianza Longitudinal

> Adapted from Templin (2018), consulted at 24th of march of 2020 from: [here](https://jonathantemplin.com/wp-content/uploads/2017/08/EPSY906_Example07b_CFA_Longitudinal_Invariance.nb_.html)

### 1. Configural Longitudinal Invariance Model (everything separate across time)

```{r configural}
configuralSyntax = "
#===================================================================================================
#Factor loadings all freely estimated in each group with labels
perc_merit1  =~ L1T1*perc_effort_w1 + L2T1*perc_talent_w1 
perc_merit2  =~ L1T2*perc_effort_w2 + L2T2*perc_talent_w2
perc_merit3  =~ L1T3*perc_effort_w3 + L2T3*perc_talent_w3

perc_nmerit1 =~ L3T1*perc_wpart_w1  + L4T1*perc_netw_w1  
perc_nmerit2 =~ L3T2*perc_wpart_w2  + L4T2*perc_netw_w2  
perc_nmerit3 =~ L3T3*perc_wpart_w3  + L4T3*perc_netw_w3

pref_merit1  =~ L5T1*pref_effort_w1 + L6T1*pref_talent_w1
pref_merit2  =~ L5T2*pref_effort_w2 + L6T2*pref_talent_w2
pref_merit3  =~ L5T3*pref_effort_w3 + L6T3*pref_talent_w3

pref_nmerit1 =~ L7T1*pref_wpart_w1 + L8T1*pref_netw_w1
pref_nmerit2 =~ L7T2*pref_wpart_w2 + L8T2*pref_netw_w2
pref_nmerit3 =~ L7T3*pref_wpart_w3 + L8T3*pref_netw_w3

#===================================================================================================
#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;

#===================================================================================================
#Residual variances all freely estimated in each group with labels 
perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
#Factor variance fixed to 1 for identification in each group
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ 1*perc_merit2
perc_merit3  ~~ 1*perc_merit3

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ 1*perc_nmerit2
perc_nmerit3 ~~ 1*perc_nmerit3

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ 1*pref_merit2
pref_merit3  ~~ 1*pref_merit3

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ 1*pref_nmerit2
pref_nmerit3 ~~ 1*pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification in each group
perc_merit1  ~ 0
perc_merit2  ~ 0
perc_merit3  ~ 0

perc_nmerit1 ~ 0
perc_nmerit2 ~ 0
perc_nmerit3 ~ 0

pref_merit1  ~ 0
pref_merit2  ~ 0
pref_merit3  ~ 0

pref_nmerit1 ~ 0
pref_nmerit2 ~ 0
pref_nmerit3 ~ 0

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#configural invariance for mode 01
conf_m1 = lavaan(model = configuralSyntax, data = wide_m1, estimator = "MLR")
```

```{r}
#configural invariance for mode 02
conf_m2 = lavaan(model = configuralSyntax, data = wide_m2, estimator = "MLR")
```

```{r}
#configural invariance for mode 03
conf_m3 = lavaan(model = configuralSyntax, data = wide_m3, estimator = "MLR")
```

```{r}
#configural invariance for complete sample (mode 1, 2 & 3)
conf = lavaan(model = configuralSyntax, data = wide02, estimator = "MLR")
```

```{r}
kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),format = "markdown",digits = 3)
```

### 2. Metric Invariance Models (ALL loadings held equal across time - identified model using Time1 Factor Variance = 1)

```{r metric}
metricSyntax1 = "
#===================================================================================================
#Factor loadings all constrained across groups with labels *****
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3

perc_nmerit1 =~ L3*perc_wpart_w1  + L4*perc_netw_w1  
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3

pref_merit1  =~ L5*pref_effort_w1 + L6*pref_talent_w1
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3

pref_nmerit1 =~ L7*pref_wpart_w1 + L8*pref_netw_w1
pref_nmerit2 =~ L7*pref_wpart_w2 + L8*pref_netw_w2
pref_nmerit3 =~ L7*pref_wpart_w3 + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;

#===================================================================================================
#Residual variances all freely estimated in each group with labels 
perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
# Factor variance fixed to 1 for identification in first group; estimated in others *****
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
perc_merit3  ~~ perc_merit3  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
perc_nmerit3 ~~ perc_nmerit3 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
pref_merit3  ~~ pref_merit3  # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ pref_nmerit2 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)
pref_nmerit3 ~~ pref_nmerit3 # deje fijado la varianza de los factores en 1 (esto lo cambie yo respecto al script original)

#===================================================================================================
#Factor mean fixed to zero for identification in each group
perc_merit1  ~ 0
perc_merit2  ~ 0
perc_merit3  ~ 0

perc_nmerit1 ~ 0
perc_nmerit2 ~ 0
perc_nmerit3 ~ 0

pref_merit1  ~ 0
pref_merit2  ~ 0
pref_merit3  ~ 0

pref_nmerit1 ~ 0
pref_nmerit2 ~ 0
pref_nmerit3 ~ 0

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#metric invariance for mode 01
metr_m1A = lavaan(model = metricSyntax1, data = wide_m1, estimator = "MLR")
```

```{r}
#metric invariance for mode 02
metr_m2A = lavaan(model = metricSyntax1, data = wide_m2, estimator = "MLR")
```

```{r}
#metric invariance for mode 03
metr_m3A = lavaan(model = metricSyntax1, data = wide_m3, estimator = "MLR")
```

```{r}
#metric invariance for complete sample (mode 1, 2 & 3)
metrA = lavaan(model = metricSyntax1, data = wide02, estimator = "MLR")
```

```{r}
#lavTestLRT(metr_m1A, conf_m1,scaled.shifted = TRUE, method = "satorra.bentler.2001")#conf_m1 was not estimated using scaled (robust) fit measures  
# lavTestLRT(metr_m2A, conf_m2,scaled.shifted = TRUE, method = "satorra.bentler.2001") # invariantes
# lavTestLRT(metr_m3A, conf_m3,scaled.shifted = TRUE, method = "satorra.bentler.2001") # invariantes

kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")

kable(lavTestLRT(metrA,conf,scaled.shifted = TRUE, method = "satorra.bentler.2001"),digits = 3,format = "markdown") # no es invariante, ver modind
```

**Mod Indices**:

```{r, modind metric1}
#get modification indices for all parameters:
metricMI1 = modificationindices(metrA, free.remove = FALSE, maximum.number = 10000)
#restrict output to only factor loading parameters
metricMI1a = metricMI1[which(metricMI1$op == "=~"),]
#restrict output to only factors and items with same time point
metricMI1a$factorTime = metricMI1a$lhs
#extract time characters from items
metricMI1a$itemTime = ifelse(test =endsWith(x = metricMI1a$rhs,suffix = "w1"),yes = "wave01",
                              no = ifelse(test = endsWith(x = metricMI1a$rhs,suffix = "w2"),yes = "wave02",
                                          no = "wave03"))

metricMI1a$factorTime = ifelse(test =endsWith(x = metricMI1a$lhs,suffix = "1"),yes = "wave01",
                               no = ifelse(test = endsWith(x = metricMI1a$lhs,suffix = "2"),yes = "wave02",
                                           no = "wave03"))
metricMI1a$factor     = ifelse(test =startsWith(x = metricMI1a$lhs,prefix = "perc_merit"),yes = "perc_merit",
                               no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "perc_nmerit"),yes = "perc_nmerit",
                                           no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "pref_merit"),yes = "pref_merit",
                                                       no = "pref_nmerit" )))

metricMI1a$item     = ifelse(test =startsWith(x = metricMI1a$rhs,prefix = "perc_effort") | startsWith(x = metricMI1a$rhs,prefix = "perc_talent"),yes = "perc_merit",
                               no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "perc_wpart")  | startsWith(x = metricMI1a$rhs,prefix = "perc_netw"),yes = "perc_nmerit",
                                           no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "pref_effort") | startsWith(x = metricMI1a$rhs,prefix = "pref_talent"),yes = "pref_merit",
                                                       no = "pref_nmerit" )))

metricMI1b<- metricMI1a %>% filter(factorTime==itemTime & factor==item) # quedarme con los que son de la misma ola y mismo factor

#show only MIs for same time
metricMI1b <- metricMI1b[order(-metricMI1b$mi),]
```

```{r metric2}
metricSyntax2 = "
#===================================================================================================
#Factor loadings all constrained across groups with labels *****
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1 #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6*pref_talent_w1
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;

#===================================================================================================
#Residual variances all freely estimated in each group with labels 
perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
# Factor variance fixed to 1 for identification in first group; estimated in others *****
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3
perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3
pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3
pref_nmerit1 ~~ 1*pref_nmerit1 
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification in each group
perc_merit1  ~ 0
perc_merit2  ~ 0
perc_merit3  ~ 0
perc_nmerit1 ~ 0
perc_nmerit2 ~ 0
perc_nmerit3 ~ 0
pref_merit1  ~ 0
pref_merit2  ~ 0
pref_merit3  ~ 0
pref_nmerit1 ~ 0
pref_nmerit2 ~ 0
pref_nmerit3 ~ 0

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3
perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3
pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3
pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
"
```

```{r}
#metric invariance (v02) for complete sample (mode 1, 2 & 3) +  free parameter
metrA2 = lavaan(model = metricSyntax2, data = wide02, estimator = "MLR")

kable(rbind(fit.conf=fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")

#fit.metrA  = all factor loading restricted
#fit.metrA2 = one free estimated factor loading
kable(lavTestLRT(metrA2,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3) #non invariant, see modind metric2 below
```

**NOTE**: El modelo que deja libre 1 parámetro (metrA2) mejora el ajuste, pero la diff del chi2 sigue siendo significativa al 0.05.

**Mod Indices**:

```{r, modind metric2}
#get modification indices for all parameters:
metricMI1 = modificationindices(metrA2, free.remove = FALSE, maximum.number = 10000)
#restrict output to only factor loading parameters
metricMI1a = metricMI1[which(metricMI1$op == "=~"),]
#restrict output to only factors and items with same time point
metricMI1a$factorTime = metricMI1a$lhs
#extract time characters from items
metricMI1a$itemTime   = ifelse(test =endsWith(x = metricMI1a$rhs,suffix = "w1"),yes = "wave01",
                                no = ifelse(test = endsWith(x = metricMI1a$rhs,suffix = "w2"),yes = "wave02",
                                            no = "wave03"))

metricMI1a$factorTime = ifelse(test =endsWith(x = metricMI1a$lhs,suffix = "1"),yes = "wave01",
                               no = ifelse(test = endsWith(x = metricMI1a$lhs,suffix = "2"),yes = "wave02",
                                           no = "wave03"))
metricMI1a$factor     = ifelse(test =startsWith(x = metricMI1a$lhs,prefix = "perc_merit"),yes = "perc_merit",
                               no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "perc_nmerit"),yes = "perc_nmerit",
                                           no = ifelse(test = startsWith(x = metricMI1a$lhs,prefix = "pref_merit"),yes = "pref_merit",
                                                       no = "pref_nmerit" )))

metricMI1a$item      = ifelse(test =startsWith(x = metricMI1a$rhs,prefix = "perc_effort") | startsWith(x = metricMI1a$rhs,prefix = "perc_talent"),yes = "perc_merit",
                              no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "perc_wpart")  | startsWith(x = metricMI1a$rhs,prefix = "perc_netw"),yes = "perc_nmerit",
                                          no = ifelse(test = startsWith(x = metricMI1a$rhs,prefix = "pref_effort") | startsWith(x = metricMI1a$rhs,prefix = "pref_talent"),yes = "pref_merit",
                                                      no = "pref_nmerit" )))

metricMI1b<- metricMI1a %>% filter(factorTime==itemTime & factor==item) # quedarme con los que son de la misma ola y mismo factor

#show only MIs for same time
metricMI1b <- metricMI1b[order(-metricMI1b$mi),]
```

```{r metric3}
metricSyntax3 = "
#===================================================================================================
#Factor loadings all constrained across groups with labels *****
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1  #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all freely estimated in each group with labels (8 items for each wave 8x3 = 16 intercepts)
perc_effort_w1 ~ I1T1*1; perc_talent_w1 ~ I2T1*1; perc_wpart_w1 ~ I3T1*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5T1*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7T1*1; pref_netw_w1 ~ I8T1*1; 
perc_effort_w2 ~ I1T2*1; perc_talent_w2 ~ I2T2*1; perc_wpart_w2 ~ I3T2*1; perc_netw_w2 ~ I4T2*1; pref_effort_w2 ~ I5T2*1; pref_talent_w2 ~ I6T2*1; pref_wpart_w2 ~ I7T2*1; pref_netw_w2 ~ I8T2*1; 
perc_effort_w3 ~ I1T3*1; perc_talent_w3 ~ I2T3*1; perc_wpart_w3 ~ I3T3*1; perc_netw_w3 ~ I4T3*1; pref_effort_w3 ~ I5T3*1; pref_talent_w3 ~ I6T3*1; pref_wpart_w3 ~ I7T3*1; pref_netw_w3 ~ I8T3*1;

#===================================================================================================
#Residual variances all freely estimated in each group with labels 
perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)
perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
# Factor variance fixed to 1 for identification in first group; estimated in others *****
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3
perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3
pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3
pref_nmerit1 ~~ 1*pref_nmerit1 
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification in each group
perc_merit1  ~ 0
perc_merit2  ~ 0
perc_merit3  ~ 0
perc_nmerit1 ~ 0
perc_nmerit2 ~ 0
perc_nmerit3 ~ 0
pref_merit1  ~ 0
pref_merit2  ~ 0
pref_merit3  ~ 0
pref_nmerit1 ~ 0
pref_nmerit2 ~ 0
pref_nmerit3 ~ 0

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3
perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3
pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3
pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
"
```

```{r}
#metric invariance (v03) for complete sample (mode 1, 2 & 3) +  free parameter
metrA3 = lavaan(model = metricSyntax3, data = wide02, estimator = "MLR")

#metrA  = all factor loading restricted
#metrA2 = one free estimated factor loading
#metrA3 = two free estimated factor loading

kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")

kable(lavTestLRT(metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
```

**NOTE**: the model with two free parameters (metric v03) at factor loadings works fine. Chi2 diff test is not significant

### 3. Scalar Invariance Models (ALL loadings held equal across time - identified model using Time1 Factor Mean = 1)

```{r, scalar1}
scalarSyntax1 = "
#===================================================================================================
#Factor loadings constrained across groups except non-invariant ones
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all constrained across groups except non-invariant ones ****
perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;

#===================================================================================================
#Residual variances all freely estimated in each group with labels 
perc_effort_w1 ~~ E1T1*perc_effort_w1; perc_talent_w1 ~~ E2T1*perc_talent_w1; perc_wpart_w1 ~~ E3T1*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5T1*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7T1*pref_wpart_w1; pref_netw_w1 ~~ E8T1*pref_netw_w1;
perc_effort_w2 ~~ E1T2*perc_effort_w2; perc_talent_w2 ~~ E2T2*perc_talent_w2; perc_wpart_w2 ~~ E3T2*perc_wpart_w2; perc_netw_w2 ~~ E4T2*perc_netw_w2; pref_effort_w2 ~~ E5T2*pref_effort_w2; pref_talent_w2 ~~ E6T2*pref_talent_w2; pref_wpart_w2 ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8T2*pref_netw_w2;
perc_effort_w3 ~~ E1T3*perc_effort_w3; perc_talent_w3 ~~ E2T3*perc_talent_w3; perc_wpart_w3 ~~ E3T3*perc_wpart_w3; perc_netw_w3 ~~ E4T3*perc_netw_w3; pref_effort_w3 ~~ E5T3*pref_effort_w3; pref_talent_w3 ~~ E6T3*pref_talent_w3; pref_wpart_w3 ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8T2*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
#Factor variance fixed to 1 for identification in first group; estimated in others
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification first group, estimated in others ****
perc_merit1  ~ 0
perc_merit2  ~ 1
perc_merit3  ~ 1

perc_nmerit1 ~ 0
perc_nmerit2 ~ 1
perc_nmerit3 ~ 1

pref_merit1  ~ 0
pref_merit2  ~ 1
pref_merit3  ~ 1

pref_nmerit1 ~ 0
pref_nmerit2 ~ 1
pref_nmerit3 ~ 1

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#scalar invariance for mode 01
sca_m1A = lavaan(model = scalarSyntax1, data = wide_m1, estimator = "MLR")
```

```{r}
#scalar invariance for mode 02
sca_m2A = lavaan(model = scalarSyntax1, data = wide_m2, estimator = "MLR")
```

```{r}
#scalar invariance for mode 03
sca_m3A = lavaan(model = scalarSyntax1, data = wide_m3, estimator = "MLR")
```

```{r}
#scalar invariance for complete sample (mode 1, 2 & 3)
scaA = lavaan(model = scalarSyntax1, data = wide02, estimator = "MLR")
```

```{r}
kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
```

```{r}
kable(lavTestLRT(scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
```

**NOTE**: the chi2 test shows that scalar vs metric models are not significant.

### 4. Residual Variance Invariance Model (error variances held equal for all except non-invariant items)

```{r residual1}
residVarSyntax1 = "
#===================================================================================================
#Factor loadings constrained across groups except non-invariant ones
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all constrained across groups except non-invariant ones 
perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;

#===================================================================================================
#Redidual variances all constrained in each group except non-invariant ones ****
perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2;
perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
#Factor variance fixed to 1 for identification in first group; estimated in others
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification first group, estimated in others ****
perc_merit1  ~ 0
perc_merit2  ~ 1
perc_merit3  ~ 1

perc_nmerit1 ~ 0
perc_nmerit2 ~ 1
perc_nmerit3 ~ 1

pref_merit1  ~ 0
pref_merit2  ~ 1
pref_merit3  ~ 1

pref_nmerit1 ~ 0
pref_nmerit2 ~ 1
pref_nmerit3 ~ 1

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#residual invariance for complete sample (mode 1, 2 & 3)
resA = lavaan(model =residVarSyntax1 , data = wide02, estimator = "MLR")
```

```{r}
kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
```

```{r}
kable(lavTestLRT(resA,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
```

**Mod Indices**:

```{r modind residual1}
#get modification indices for all parameters:
residVarMI1 = modificationindices(resA, free.remove = FALSE, maximum.number = 10000)
#restrict output to only residual variance parameters:
residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
residVarMI1a = residVarMI1a[order(-residVarMI1a$mi),] # change: pref_wpart_w2 ~~  pref_wpart_w2
```

```{r residual2}
residVarSyntax2 = "
#===================================================================================================
#Factor loadings constrained across groups except non-invariant ones
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all constrained across groups except non-invariant ones 
perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;

#===================================================================================================
#Redidual variances all constrained in each group except non-invariant ones ****
perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2; # F-RES
perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3;
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
#Factor variance fixed to 1 for identification in first group; estimated in others
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification first group, estimated in others ****
perc_merit1  ~ 0
perc_merit2  ~ 1
perc_merit3  ~ 1

perc_nmerit1 ~ 0
perc_nmerit2 ~ 1
perc_nmerit3 ~ 1

pref_merit1  ~ 0
pref_merit2  ~ 1
pref_merit3  ~ 1

pref_nmerit1 ~ 0
pref_nmerit2 ~ 1
pref_nmerit3 ~ 1

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#residual invariance (v02) for complete sample (mode 1, 2 & 3)
resA2 = lavaan(model =residVarSyntax2 , data = wide02, estimator = "MLR")
```

```{r}
kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA2  =fitmeasures(resA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")

```

```{r}
kable(lavTestLRT(resA2,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
```

**Mod Indices**:

```{r modind residual2}
#get modification indices for all parameters:
residVarMI1 = modificationindices(resA2, free.remove = FALSE, maximum.number = 10000)
#restrict output to only residual variance parameters:
residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
residVarMI1a <- residVarMI1a[order(-residVarMI1a$mi),] # changes: pref_wpart_w3 ~~  pref_wpart_w3
```

```{r residual3}
residVarSyntax3 = "
#===================================================================================================
#Factor loadings constrained across groups except non-invariant ones
perc_merit1  =~ L1*perc_effort_w1 + L2*perc_talent_w1 
perc_merit2  =~ L1*perc_effort_w2 + L2*perc_talent_w2
perc_merit3  =~ L1*perc_effort_w3 + L2*perc_talent_w3
perc_nmerit1 =~ L3*perc_wpart_w1  + L4T1*perc_netw_w1   #FREE PARAMETER    
perc_nmerit2 =~ L3*perc_wpart_w2  + L4*perc_netw_w2  
perc_nmerit3 =~ L3*perc_wpart_w3  + L4*perc_netw_w3    
pref_merit1  =~ L5*pref_effort_w1 + L6T1*pref_talent_w1 #FREE PARAMETER
pref_merit2  =~ L5*pref_effort_w2 + L6*pref_talent_w2
pref_merit3  =~ L5*pref_effort_w3 + L6*pref_talent_w3
pref_nmerit1 =~ L7*pref_wpart_w1  + L8*pref_netw_w1    
pref_nmerit2 =~ L7*pref_wpart_w2  + L8*pref_netw_w2  
pref_nmerit3 =~ L7*pref_wpart_w3  + L8*pref_netw_w3

#===================================================================================================
#Item intercepts all constrained across groups except non-invariant ones 
perc_effort_w1 ~ I1*1; perc_talent_w1 ~ I2*1; perc_wpart_w1 ~ I3*1; perc_netw_w1 ~ I4T1*1; pref_effort_w1 ~ I5*1; pref_talent_w1 ~ I6T1*1; pref_wpart_w1 ~ I7*1; pref_netw_w1 ~ I8*1; # TWO FREE INTERCEPTS
perc_effort_w2 ~ I1*1; perc_talent_w2 ~ I2*1; perc_wpart_w2 ~ I3*1; perc_netw_w2 ~ I4*1; pref_effort_w2   ~ I5*1; pref_talent_w2 ~ I6*1; pref_wpart_w2   ~ I7*1; pref_netw_w2 ~ I8*1; 
perc_effort_w3 ~ I1*1; perc_talent_w3 ~ I2*1; perc_wpart_w3 ~ I3*1; perc_netw_w3 ~ I4*1; pref_effort_w3   ~ I5*1; pref_talent_w3 ~ I6*1; pref_wpart_w3   ~ I7*1; pref_netw_w3 ~ I8*1;

#===================================================================================================
#Redidual variances all constrained in each group except non-invariant ones ****
perc_effort_w1 ~~ E1*perc_effort_w1; perc_talent_w1 ~~ E2*perc_talent_w1; perc_wpart_w1 ~~ E3*perc_wpart_w1; perc_netw_w1 ~~ E4T1*perc_netw_w1; pref_effort_w1 ~~ E5*pref_effort_w1; pref_talent_w1 ~~ E6T1*pref_talent_w1; pref_wpart_w1 ~~ E7*pref_wpart_w1; pref_netw_w1 ~~ E8*pref_netw_w1;
perc_effort_w2 ~~ E1*perc_effort_w2; perc_talent_w2 ~~ E2*perc_talent_w2; perc_wpart_w2 ~~ E3*perc_wpart_w2; perc_netw_w2 ~~ E4*perc_netw_w2; pref_effort_w2   ~~ E5*pref_effort_w2; pref_talent_w2 ~~ E6*pref_talent_w2; pref_wpart_w2   ~~ E7T2*pref_wpart_w2; pref_netw_w2 ~~ E8*pref_netw_w2; # F-RES2
perc_effort_w3 ~~ E1*perc_effort_w3; perc_talent_w3 ~~ E2*perc_talent_w3; perc_wpart_w3 ~~ E3*perc_wpart_w3; perc_netw_w3 ~~ E4*perc_netw_w3; pref_effort_w3   ~~ E5*pref_effort_w3; pref_talent_w3 ~~ E6*pref_talent_w3; pref_wpart_w3   ~~ E7T3*pref_wpart_w3; pref_netw_w3 ~~ E8*pref_netw_w3; # F-RES3
#===================================================================================================
#Residual covariances freely estimated in each group with labels (C1T12= covariance 1 between time 1&2)

perc_effort_w1 ~~ C1T12*perc_effort_w2 + C1T13*perc_effort_w3; perc_effort_w2 ~~ C1T23*perc_effort_w3;  
perc_talent_w1 ~~ C2T12*perc_talent_w2 + C2T13*perc_talent_w3; perc_talent_w2 ~~ C2T23*perc_talent_w3;  
perc_wpart_w1  ~~ C3T12*perc_wpart_w2  + C3T13*perc_wpart_w3 ; perc_wpart_w2  ~~ C3T23*perc_wpart_w3 ; 
perc_netw_w1   ~~ C4T12*perc_netw_w2   + C4T13*perc_netw_w3  ; perc_netw_w2   ~~ C4T23*perc_netw_w3  ;
pref_effort_w1 ~~ C5T12*pref_effort_w2 + C5T13*pref_effort_w3; pref_effort_w2 ~~ C5T23*pref_effort_w3;  
pref_talent_w1 ~~ C6T12*pref_talent_w2 + C6T13*pref_talent_w3; pref_talent_w2 ~~ C6T23*pref_talent_w3;  
pref_wpart_w1  ~~ C7T12*pref_wpart_w2  + C7T13*pref_wpart_w3 ; pref_wpart_w2  ~~ C7T23*pref_wpart_w3 ; 
pref_netw_w1   ~~ C8T12*pref_netw_w2   + C8T13*pref_netw_w3  ; pref_netw_w2   ~~ C8T23*pref_netw_w3  ;

#===================================================================================================
#Factor variance fixed to 1 for identification in first group; estimated in others
perc_merit1  ~~ 1*perc_merit1
perc_merit2  ~~ perc_merit2
perc_merit3  ~~ perc_merit3

perc_nmerit1 ~~ 1*perc_nmerit1
perc_nmerit2 ~~ perc_nmerit2
perc_nmerit3 ~~ perc_nmerit3

pref_merit1  ~~ 1*pref_merit1
pref_merit2  ~~ pref_merit2
pref_merit3  ~~ pref_merit3

pref_nmerit1 ~~ 1*pref_nmerit1
pref_nmerit2 ~~ pref_nmerit2
pref_nmerit3 ~~ pref_nmerit3

#===================================================================================================
#Factor mean fixed to zero for identification first group, estimated in others ****
perc_merit1  ~ 0
perc_merit2  ~ 1
perc_merit3  ~ 1

perc_nmerit1 ~ 0
perc_nmerit2 ~ 1
perc_nmerit3 ~ 1

pref_merit1  ~ 0
pref_merit2  ~ 1
pref_merit3  ~ 1

pref_nmerit1 ~ 0
pref_nmerit2 ~ 1
pref_nmerit3 ~ 1

#===================================================================================================
#Factor covariances all freely estimated
perc_merit1  ~~ perc_merit2 + perc_merit3
perc_merit2  ~~ perc_merit3

perc_nmerit1 ~~ perc_nmerit2 + perc_nmerit3
perc_nmerit2 ~~ perc_nmerit3

pref_merit1  ~~ pref_merit2 + pref_merit3
pref_merit2  ~~ pref_merit3

pref_nmerit1 ~~ pref_nmerit2 + pref_nmerit3
pref_nmerit2 ~~ pref_nmerit3
#===================================================================================================
"
```

```{r}
#residual invariance (v03) for complete sample (mode 1, 2 & 3)
resA3 = lavaan(model =residVarSyntax3 , data = wide02, estimator = "MLR")
```

```{r}
kable(rbind(fit.conf   =fitmeasures(conf)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA  =fitmeasures(metrA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA2 =fitmeasures(metrA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.metrA3 =fitmeasures(metrA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.scaA   =fitmeasures(scaA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA   =fitmeasures(resA)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA2  =fitmeasures(resA2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")],
            fit.resA3  =fitmeasures(resA3)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")]),digits = 3,format = "markdown")
```

```{r}
kable(lavTestLRT(resA3,scaA,metrA3,conf,scaled.shifted = TRUE,method = "satorra.bentler.2001"),format = "markdown",digits = 3)
```

**Mod Indices**:

```{r modind residual3}
#get modification indices for all parameters:
residVarMI1 = modificationindices(resA3, free.remove = FALSE, maximum.number = 10000)
#restrict output to only residual variance parameters:
residVarMI1a = residVarMI1[which(residVarMI1$op == "~~" & residVarMI1$lhs == residVarMI1$rhs),]
residVarMI1a <- residVarMI1a[order(-residVarMI1a$mi),] # change: pref_wpart_w3 ~~  pref_wpart_w3 (?)
```

**NOTE**: The residual is too big. Mod indices suggest to freely estimate the pref_wpart_w3 covariance, but this one is already free.

* Invarianza Longitudinal en lavaan
  * https://quantdev.ssri.psu.edu/tutorials/intro-basics-longitudinal-measurement-invariance 
  * https://jonathantemplin.com/wp-content/uploads/2017/08/EPSY906_Example07b_CFA_Longitudinal_Invariance.nb_.html

* https://users.ugent.be/~yrosseel/lavaan/multiplegroup6Dec2012.pdf

* Invarianza con Indicatores categóricos
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5121102/pdf/nihms757737.pdf

* Parametrización tetha: 
https://groups.google.com/forum/#!topic/lavaan/nfdatPgLLhc

* Colapsing categories:
https://www.tandfonline.com/doi/abs/10.1080/10705511.2018.1547640?journalCode=hsem20