### 2020-04-10 production/prod_prep-data-A.rmd

---
title: "Data preparation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: "../input/bib/style.css"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = ".")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

load packages
```{r}
library(sjPlot)
library(dplyr)
library(tidyr)
library(lavaan)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(summarytools)
library(sjlabelled)
library(Hmisc)
```

# Data preparation according order of the items

```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
```

##  homogenize names of the items on the scale.

The items on the scale had different orders according to the order in which they were exposed to the respondents. There are three order. The first one the orders keeps together the items of the same factor. The second order then asks for perceptions and preferences regarding the same topic (i.e Distribution of resources according to efforts). In the third order, the items are presented randomly to each respondent.

```{r}
dat04 <- w01 %>% filter(Intro==1) %>%
  select(starts_with("meritv01"),
         starts_with("meritv02"),
         starts_with("meritv03_p"),
         starts_with("FL_21_DO"))
dat04$grupo <- NA
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
dat04$grupo[dat04$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
dat04$perc_effort <- rowSums(dat04[,c(matches(match = "perc_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_talent <- rowSums(dat04[,c(matches(match = "perc_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$perc_wpart  <- rowSums(dat04[,c(matches(match = "perc_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$perc_netw   <- rowSums(dat04[,c(matches(match = "perc_netw"  ,vars = names(dat04)))],na.rm = TRUE)
dat04$pref_effort <- rowSums(dat04[,c(matches(match = "pref_effort",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_talent <- rowSums(dat04[,c(matches(match = "pref_talent",vars = names(dat04)))],na.rm = TRUE)
dat04$pref_wpart  <- rowSums(dat04[,c(matches(match = "pref_wpart" ,vars = names(dat04)))],na.rm = TRUE)
dat04$pref_netw   <- rowSums(dat04[,c(matches(match = "pref_netw"  ,vars = names(dat04)))],na.rm = TRUE)
dat04[dat04==0] <- NA
dat04 <- dat04 %>% select(starts_with(match = "perc_"),starts_with(match = "pref_"),grupo) %>% na.omit()
```

## Translation of variable labels and value labels.
```{r}
label(dat04$grupo) <- "This variable classifies the interviewees according to the order in which the indicators are shown within the scale. There are three orders  of items on the scale. In the first order (group 1), questions of the same factor are then presented to them, in the second order (group 2) they were separated and in the third order (group 3) at random."
dat04$perc_effort <- set_labels(dat04$perc_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they manage to get bigger
rewards that those who strive
less."
dat04$perc_talent <- set_labels(dat04$perc_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they manage to obtain
greater rewards than those who possess
less talent."
dat04$perc_wpart <- set_labels(dat04$perc_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have rich parents manage to get out
ahead."
dat04$perc_netw <- set_labels(dat04$perc_netw,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$perc_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
they have good contacts they manage to get out
ahead."
dat04$pref_effort <- set_labels(dat04$pref_effort,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_effort) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
the more they try they should get
greater rewards than those who
they try less."
dat04$pref_talent <- set_labels(dat04$pref_talent,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_talent) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Who
possess more talent they should get
greater rewards than those who possess
less talent."
dat04$pref_wpart <- set_labels(dat04$pref_wpart,  labels = c("strongly disagree" = 1, "disagreement" = 2, "Neither agree nor disagree"=3,"Agree"=4, "totally agree" = 5))
label(dat04$pref_wpart) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."
dat04$pref_netw <- set_labels(dat04$pref_netw,  labels = c("strongly disagree", "disagreement", "Neither agree nor disagree","Agree", "totally agree"))
label(dat04$pref_netw) <- "Thinking of Chilean society,
what measure do you find of
agree or disagree with each of
the following statements? - Is
well that those who have good
contacts get ahead."
```

```{r}
dat04 <-  dat04 %>% select(grupo, perc_effort, perc_talent, perc_wpart, perc_netw, pref_effort, pref_talent, pref_wpart , pref_netw) %>% as.data.frame()
```

```{r}
print(dfSummary(dat04, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation of group 1: order according to the factors.

```{r}
dat01 <- dat04 %>% filter(grupo==1) %>% select(-grupo) %>%  na.omit() 
```

```{r}
print(dfSummary(dat01, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation of group 2: order according to topics (i.e Effort).

```{r}
dat02 <- dat04 %>% filter(grupo==2) %>% select(perc_effort, pref_effort, perc_talent,pref_talent, perc_wpart,pref_wpart, perc_netw, pref_netw) %>% na.omit()
```

```{r}
print(dfSummary(dat02, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation of group 3: randomized order.

```{r}
dat03 <- dat04 %>% filter(grupo==3) %>% select(-grupo) %>%  na.omit()
```

```{r}
print(dfSummary(dat03, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## save data
```{r}
save(dat01,file = "input/data/proc/dat01.RData")
save(dat02,file = "input/data/proc/dat02.RData")
save(dat03,file = "input/data/proc/dat03.RData")
save(dat04,file = "input/data/proc/dat04.RData")
```

# Longitudinal data

load data
```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
w02 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola2.sav",verbose = F)
w03 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
```

## Preparation wave 1

Add tags wave 1
```{r }
label(w01$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
w01$Intro <- set_labels(w01$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 10}
w01_Intro <- w01 %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w01_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w01b <- w01 %>% filter(Intro==1, Finished==1) %>% as.data.frame()
w01b$ID<- stringr::str_split_fixed(w01b$ticket,"_", 4)[,1] # En la variable ticket el primer código es el ID, todo lo demás se borra
```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 20 }
w01b$Intro <- set_labels(w01b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w01_Introb <- w01b %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w01_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

```{r}
w01c <- w01b
w01c$grupo <- NA
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c[w01c==0] <- NA
w01c <- w01c %>% select(grupo,
                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w01c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation wave 2

Add tags wave 2
```{r }
label(w02$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
w02$Intro <- set_labels(w02$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
```

Frequency of respondents who Accept and Finish wave 2.
```{r}
w02_Intro <- w02 %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w02_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w02b <- w02 %>% filter(Intro==1, Finished==1) %>% as.data.frame()
w02b$ID <- stringr::str_split_fixed(w02b$ticket, "_", 4)[,1]
```

Frequency of respondents who Accept and Finish wave 2.
```{r }
w02b$Intro <- set_labels(w02b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w02_Introb <- w02b %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w02_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

```{r}
w02c <- w02b
w02c$treat <- NA
w02c$treat[w02c$FL_6_DO_Control==1] <- 1
w02c$treat[w02c$FL_6_DO_Tratamiento==1] <- 2
w02c$treat[w02c$FL_6_DO_Tratamiento_desigualdad==1] <- 3
w02c$treat <- factor(w02c$treat,levels = c(1,2,3),labels = c("control","pobreza","desiguald"))
w02c <- w02c %>% rename(perc_effort_w2=meritv03_perc_effort,
                        perc_talent_w2=meritv03_perc_talent,
                        perc_wpart_w2 =meritv03_perc_wpart,
                        perc_netw_w2  =meritv03_perc_netw,
                        pref_effort_w2=meritv03_pref_effort,
                        pref_talent_w2=meritv03_pref_talent,
                        pref_wpart_w2 =meritv03_pref_wpart,
                        pref_netw_w2  =meritv03_pref_netw)
w02c <- w02c %>% select(perc_effort_w2,perc_talent_w2,perc_wpart_w2,perc_netw_w2,
                        pref_effort_w2,pref_talent_w2,pref_wpart_w2,pref_netw_w2,
                        sexo_w2=sexo,edad_w2=edad,edcep_w2=edcep, treat, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w02c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation wave 3

```{r}
table(w03$Finished==1)
w03b <- filter(w03,Finished==1) # Elimina los que no aceptan
dim(w03b)
w03b$ID <- stringr::str_split_fixed(w03b$ticket, "_", 4)[,1]
```

```{r}
w03c <- w03b
w03c <- w03c %>% rename(perc_effort_w3=meritv03_perc_effort,
                        perc_talent_w3=meritv03_perc_talent,
                        perc_wpart_w3 =meritv03_perc_wpart,
                        perc_netw_w3  =meritv03_perc_netw,
                        pref_effort_w3=meritv03_pref_effort,
                        pref_talent_w3=meritv03_pref_talent,
                        pref_wpart_w3 =meritv03_pref_wpart,
                        pref_netw_w3  =meritv03_pref_netw)
w03c <- w03c %>% select(perc_effort_w3,perc_talent_w3,perc_wpart_w3,perc_netw_w3,
                        pref_effort_w3,pref_talent_w3,pref_wpart_w3,pref_netw_w3,
                        sexo_w3=sexo,edad_w3=edad,edcep_w3=edcep, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w03c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Merge wide data

```{r}
wide01 <- inner_join(x = w01c,y = w02c, by = "ID") # Solamente nos quedamos con casos en ambas olas
dim(wide01) #en w01 y w02
```

```{r}
wide02 <- inner_join(x = wide01,y = w03c, by = "ID") # Solamente nos quedamos con casos en ambas olas
dim(wide02) #en w01, w02 y w03
```

```{r}
print(dfSummary(w02, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

```{r}
save(wide01,file = "input/data/proc/datawide.RData")
save(wide02,file = "input/data/proc/datawide_123.RData")
```

# Wide to long data

```{r}
library(datasets)
library(data.table)
wide <- setDT(wide02)
long01<- melt(wide,variable.name = "wave",
              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
                                       "^sexo_","^edad_","^edcep_"),
              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
                             "pref_effort","pref_talent","pref_wpart","pref_netw",
                             "sexo","edad","edcep"))
```

```{r}
save(long01,file = "input/data/proc/datalong.RData")
```