### 2020-12-21 production/prod_analysis-group_invariance.rmd
save-files

---
title: "Group invariance "
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: "../input/bib/style.css"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: "hide"
editor_options: 
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prod_analysis-group_invariance.Rmd",output_format = "html_document"); browseURL(url = "production/")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

# Analysis

**load libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(semTools)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)

```

**load data**
```{r}
load(file = "input/data/proc/pov01_fact02.RData")
names(pov_fs)
```

```{r}
pov01 <- pov_fs %>% filter(dataset=="pvw01")
fs01  <- pov_fs %>% filter(dataset=="fsw02")
```

Meassurement model

```{r}
model01 <- '
perc_merit  =~ perc_effort + perc_talent
perc_nmerit =~ perc_wpart  + perc_netw
pref_merit  =~ pref_effort + pref_talent
pref_nmerit =~ pref_wpart  + pref_netw'
```

## Poverty Experiment Sample (wave 01)
```{r}
pov1_c <- cfa(model = model01,data = pov01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust

pov1_o <- cfa(model = model01,data = pov01,ordered = c("perc_effort","perc_talent",
                                                       "perc_wpart","perc_netw",
                                                       "pref_effort","pref_talent",
                                                       "pref_wpart","pref_netw"),std.lv=FALSE)
#---------------------------------------------------#
cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
kable(left_join(x = standardizedsolution(pov1_c) %>% 
                  filter(op=="=~") %>% 
                  select(lhs,rhs,est.std),y = standardizedsolution(pov1_o) %>% 
                  filter(op=="=~") %>%
                  select(lhs,rhs,est.std),c("lhs","rhs")),
      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
```

```{r eval=FALSE, include=FALSE}
fitmeasures(pov1_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(pov1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

## Factorial Survey sample (wave 02)

```{r}
fs1_c <- cfa(model = model01,data = fs01,estimator="MLR",std.lv=FALSE) # Continuous/ estimator ML Robust

fs1_o <- cfa(model = model01,data = fs01,ordered = c("perc_effort","perc_talent",
                                                      "perc_wpart","perc_netw",
                                                      "pref_effort","pref_talent",
                                                      "pref_wpart","pref_netw"),std.lv=FALSE)
#---------------------------------------------------#
cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
kable(left_join(x = standardizedsolution(fs1_c) %>% 
                  filter(op=="=~") %>% 
                  select(lhs,rhs,est.std),y = standardizedsolution(fs1_o) %>% 
                  filter(op=="=~") %>%
                  select(lhs,rhs,est.std),c("lhs","rhs")),
      format = "markdown",digits = 2,col.names = cnames, caption = "Factor loadings")
```

```{r eval=FALSE, include=FALSE}
fitmeasures(fs1_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

## Group CFA
```{r}
group_c <- cfa(model01, 
           data = pov_fs, 
           group = "dataset",estimator="MLR",std.lv=FALSE)

group_o <- cfa(model = model01,data = pov_fs,ordered = c("perc_effort","perc_talent",
                                                        "perc_wpart","perc_netw",
                                                        "pref_effort","pref_talent",
                                                        "pref_wpart","pref_netw"),std.lv=FALSE,group = "dataset")
```

```{r}
summary(group_c, standardized=TRUE,fit.measures=TRUE)
summary(group_o, standardized=TRUE,fit.measures=TRUE)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(group_c, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
fitmeasures(group_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]
```

# Tables
## CFA

```{r}
labs<- c("A. Who the more they try they manage to get bigger rewards that those who striveless.",
         "B. Who possess more talent they manage to obtain greater rewards than those who possess less talent.",
         "C. Who they have rich parents manage to get out ahead.",
         "D. Who they have good contacts they manage to get out ahead.",
         "E. Who the more they try they should get greater rewards than those who they try less." ,
         "F. Who possess more talent they should get greater rewards than those who possess less talent." ,
         "G. It's fine that those who have rich parents get ahead",
         "H. Is well that those who have good contacts get ahead." )
```

```{r}
loads01<- data.frame(round(inspect(fs1_o,what="std")$lambda,digits = 2))
loads01[loads01==c(0.00)] <- NA; names(loads01) <- c(paste0("M1.",1:4))

for (i in names(loads01)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  loads01[,i] <- sprintf(loads01[,i], fmt = '%#.2f')
}
loads01[loads01=="NA"] <- ""
```

```{r tb_loads-fit-fac02}
fm01<- data.frame(t(data.frame(fitmeasures(fs1_o, output ="matrix")[c("chisq.scaled","df","cfi.scaled","rmsea.scaled"),]))); row.names(fm01) ="M1.1"

#------chi2, df------#
fm04.1 <- fm01 %>% select(chisq.scaled,df) 
fm04.1$chisq.scaled <- round(x = fm04.1$chisq.scaled,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq.scaled,"(",fm04.1$df,")")
fm04.1 <- select(fm04.1,"chi2df")

#------CFI, RMSEA------#
fm04.2 <- fm01 %>% select(cfi.scaled,rmsea.scaled)
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

#------ merge fit indices ------#
fm.df      <- bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(nobs(fs1_o)) 
fm.df      <- data.frame(t(fm.df))

#------ merge ------#
tb.fm<- bind_rows(loads01,fm.df)
tb.fm<- tb.fm %>% 
  mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{RMSEA}$","$N$")) %>%
  select(vars,everything())
#--table---#
tb.col  <- c("Variables",1:4)
tb.foot <- paste0("; p<0.05=$^{*}$; p<0.01=$^{**}$; p<0.001=$^{***}$. Standardised factor loadings using DWLS estimator. CFI = Comparative fit index (scaled), RMSEA = Root mean square error of approximation (scaled)")
tb.caption <- "Factor loadings and fit measures"

tb.fm01<- kable(tb.fm,escape = FALSE,align = "lcccc",col.names = tb.col, caption = tb.caption) %>% 
          kable_classic(full_width = F,font_size = 9,bootstrap_options = c("striped", "hover"),html_font = "Sans") %>% 
          add_header_above(header = c(" "=1,"Factor loadings"= 4)) %>%
          column_spec(column = 1,width = "5cm") %>% 
          column_spec(column = c(2:5),width = "0.5cm") %>% 
          row_spec(row = 8,hline_after = TRUE) %>%
          add_indent(c(9:12)) %>% 
          footnote(general =tb.foot ,footnote_as_chunk = T);tb.fm01

save(tb.fm,file = "output/tables/tb.fact-f02.RData")
```

## Invariance

```{r}
inv01<- measurementInvariance(model=model01,data=pov_fs,group="dataset",estimator = "ML",strict=TRUE)
```

```{r}
conf  <- inv01$fit.configural
weak  <- inv01$fit.loadings
strong<- inv01$fit.intercepts
strict<- inv01$fit.residuals

tab01<- anova(conf,weak,strong,strict,SB.classic=TRUE) %>% 
  as_tibble() %>% 
  select("Chisq","Df","chisq_diff"=`Chisq diff`,"df_diff"=`Df diff`,"pvalue"=`Pr(>Chisq)`) %>% 
  mutate(stars=gtools::stars.pval(pvalue),
         chisqt=paste0(round(Chisq,2)," (",Df,") "),
         decision=ifelse(pvalue>0.05,yes = "Accept",no = "Reject"),
         model=c("Configural","Weak","Strong","Strict"))

fit.meas<- bind_rows(fitmeasures(inv01$fit.configural,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.loadings,  output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.intercepts,output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),],
                     fitmeasures(inv01$fit.residuals, output ="matrix")[c("chisq","df","cfi","rmsea","rmsea.ci.lower","rmsea.ci.upper"),])

# compute differences in chisq, df, cfi and rmsea (90%, lower.ci - upper.ci )
fit.meas<- fit.meas %>% 
  mutate(diff.chi2 = chisq    - lag(chisq,default = first(chisq)),
         diff.df   = df       - lag(df,   default = first(df)),
         diff.cfi  = cfi      - lag(cfi,  default = first(cfi)),
         diff.rmsea   = rmsea - lag(rmsea,default = first(rmsea))) %>%
  round(3) %>% 
  mutate(rmsea.ci=paste0(rmsea," \n ", "(",rmsea.ci.lower,"-",rmsea.ci.upper,")"))

tab.inv<- bind_cols(tab01,fit.meas) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.df,diff.cfi,diff.rmsea,stars,decision) %>% 
  mutate(diff.chi2=paste0(diff.chi2," (",diff.df,") ",stars)) %>% 
  select(model,chisqt,cfi,rmsea.ci,diff.chi2,diff.cfi,diff.rmsea,decision)

#clean values
tab.inv[tab.inv == c("0 (0) ")] <- NA
tab.inv[tab.inv == c(0)] <- NA
```

```{r}
col.nam <- c("Model","$\\chi^2 (\\text{df})$","CFI","RMSEA (90% CI)",
             "$\\Delta \\chi^2 (\\Delta \\text{df}$)","$\\Delta \\text{CFI}$","$\\Delta \\text{RMSEA}$","Decision")
footnote <- paste0("$N$ = ",nobs(conf),"; group 1 $n$ = ",conf@Data@nobs[[1]],"; group 2 $n$ = ",conf@Data@nobs[[2]], "\n ***p < .001")

kable(tab.inv, col.names = col.nam, booktabs=TRUE) %>% 
  kable_classic(full_width = F, bootstrap_options = c("striped", "hover"),html_font = "Serif") %>% 
  footnote(general = footnote, footnote_as_chunk = T)

save(tab.inv,file = "output/tables/tb.inv-group.RData")
```

```{r eval=FALSE, include=FALSE}
### Check an alternative method
configural        <- cfa(model01, data=pov_fs, group = "dataset", estimator="ML")
weak.invariance   <- cfa(model01, data=pov_fs, group = "dataset",group.equal = "loadings", estimator="ML")
strong.invariance <- cfa(model01, data=pov_fs, group = "dataset",group.equal = c( "loadings", "intercepts"), estimator="ML")
strict.invariance <- cfa(model01, data=pov_fs, group = "dataset",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
```

```{r eval=FALSE, include=FALSE}
anova(configural,weak.invariance,strong.invariance,strict.invariance)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(configural       )
fitmeasures(weak.invariance  )
fitmeasures(strong.invariance) 
fitmeasures(strict.invariance) 
```

* The results are the same in both methods

* Tabla de CFA con la segunda muestra
* Tabla resultados de invarianza entre muestras  
* Modelo de reporte: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5145197/