### 2021-04-01 production/prod_analysis-cfa-study01.rmd

---
title: "Analysis PPM-S"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

**Libraries**
```{r}
library(sjPlot)
library(dplyr)
library(lavaan)
library(semPlot)
library(stargazer)
library(corrplot)
library(psych)
library(knitr)
library(kableExtra)
library(rvest)
library(sjlabelled)
library(ggplot2)
library(stargazer)
```

**Load data**

```{r}
load(file = "input/data/proc/getahead.rda")
load(file = "input/data/proc/detearn.rda")
```

```{r}
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

```

# Perception

## Descriptives
```{r}
# Summary table
names(getahead)

stargazer(getahead[, -c(1, 2, 11:18)], type = "text")

# Simple freq graph
plot_frq(getahead$wealthy,
        show.values = TRUE,
        coord.flip = TRUE)

# Likert graph
plot_likert(getahead %>% select(wealthy:gender)) # ajustar

# short var labels
items <- c("Wealthy family", "Educated parents","Ambition", 
           "Hard work", "Race","Gender", "Know people", "Pol. connect.")
# cambiar orden para la leyenda
labels = c("Essential",
           "Very important",
           "Not very important",
           "Not important at all",
           "(neutral)")

# ajustes globales sjplot

set_theme(
  theme = "blank",
  legend.inside = TRUE,
  legend.pos = "bottom right",
  legend.size = 0.8,
  geom.outline.size = 0,
  geom.label.size = 3)

# Gráfico final

plot_likert(getahead%>% select(wealthy:gender),
           axis.labels   = items,
           legend.labels = labels,
           cat.neutral   = 3,
           # identifica a indiferentes
           geom.colors   = c("#9ecae1", "#6baed6", "#4292c6",
                             "#2171b5"),
           # colorbrewer2.org para temas
           sort.frq      = "neg.asc",
           # sort descending)
           title         = "How important is for getting ahead ...",
           intercept.line.color = "white",
           # middle line white
           expand.grid   = FALSE,
           # no inner margins in plot
           show.n        = FALSE,
           # hide N's in axis labels
           grid.range    = 1.3
) # expand axis from 130% to 130%

# Save
# dev.copy(png, "./results/likert_ga.png")
# dev.off()

# Correlations

# Matrices & plot

cormat_ga = cor(getahead %>% select(wealthy:gender),
                use = "pairwise.complete.obs")
corrplot.mixed(cormat_ga)
# dev.copy(png,
#          "./results/corplot_ga.png",
#          width = 600,
#          height = 600)
# dev.off()

# Correlation table a latex
# sink("./results/cortab_ga.txt")
# cor2latex(
#   getahead[2:10],
#   stars   = TRUE,
#   adjust = "none",
#   apa     = TRUE,
#   heading = " ",
#   caption = "Reasons to get ahead")
# sink()
```

## Measurement models

### Continuos 
```{r}
# Estimación del modelo de 1er order
cfa_1 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
	  '
fit_1 <- cfa(cfa_1, data = getahead, missing = "ML")
summary(fit_1, fit.measures = TRUE, standardized = TRUE)

# Path model
semPaths(
  fit_1,
  layout = "tree2",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = TRUE
)
```

### Categorical

```{r}
# Estimación categorical 3 levels (5 levels convergence prob.)
cfa_1c <- '
		  # latent variables
		  merit=~ c3hwork + c3ambition
		  parent=~ c3wealthy + c3pareduc
		  backg=~ c3race + c3gender
			network=~ people + polcone
		  '
fit_1c <- cfa(
  cfa_1c,
  data = getahead,
  ordered = c(
    "c3hwork",
    "c3ambition",
    "c3wealthy",
    "c3pareduc",
    "c3race",
    "c3gender",
    "c3people",
    "c3polcone"
  )
)
summary(fit_1c, fit.measures = TRUE, standardized = TRUE) # works
mod = modificationindices(fit_1c)
subset(mod, mi > 3.84, select = lhs:mi)
```

```{r eval=FALSE, include=FALSE}
# Estimación del modelo 2nd order

cfa_2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit =~parent + backg + network
	  '
fit_2 <- cfa(cfa_2, data = getahead, missing = "ML")
summary(fit_2, fit.measures = TRUE, standardized = TRUE)

# Path
semPaths(
  fit_2,
  layout = "tree",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = FALSE
)
dev.copy(png,
         "./results/path_cfa2.png",
         width = 750,
         height = 600)
dev.off()
```

# Preference

## Descriprtives
```{r}    
# 2.2 Descriptives ---------------------- ######

load(file = "input/data/proc/detearn.rda")

# Summary table
names(detearn)
stargazer(detearn[,-c(1,2)], type = "text")

# Summary table to latex
stargazer(detearn[, -c(1, 2)])

# Simple freq graph
sjp.frq(detearn2$respons,
        show.values = TRUE,
        coord.flip = TRUE)

# Likert graph (uses detearn2 object, with original coding)

sjp.likert((detearn2[, -c(1, 2)])) # ajustar

# short var labels
items <- c("Responsibility",
           "Years education",
           "Family",
           "Children",
           "Well job",
           "Hard job")

# cambiar orden para la leyenda
labels = c("Essential",
           "Very important",
           "(neutral)"
           ,
           "Not very important",
           "Not important at all")

# ajustes globales sjplot
theme_set(
  theme = "blank",
  legend.inside = TRUE,
  legend.pos = "bottom right",
  legend.size = 0.8,
  geom.outline.size = 0,
  geom.label.size = 3
)

plot_likert((detearn2[, -c(1, 2)]),
           axis.labels   = items,
           legend.labels = labels,
           cat.neutral   = 3,# identifica a indiferentes
           geom.colors   = c("#9ecae1", "#6baed6", "#4292c6","#2171b5"),# colorbrewer2.org para temas
           sort.frq      = "neg.asc",# sort descending)
           title         = "In deciding how much people ought to earn, how important should each of these things be...",
           intercept.line.color = "white",# middle line white
           expand.grid   = FALSE,# no inner margins in plot
           show.n        = FALSE,# hide N's in axis labels
           grid.range    = 1.3
) # expand axis from 130% to 130%

# Save
dev.copy(png, "./results/likert_detearn.png")
dev.off()

# freq(detearn2$pareduc)   # check educated parents, ok

# Correlations
cormat = cor(detearn[5:8], use = "pairwise.complete.obs")
corrplot.mixed(cormat)

#save
dev.copy(png,
         "./results/corplot_detearn.png",
         width = 600,
         height = 600)
dev.off()
```

## Measurement models

```{r}
cfa_1 <- '
	  # latent variables
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
fit_1 <- cfa(cfa_1, data = detearn, missing = "ML")

summary(fit_1, fit.measures = TRUE, standardized = TRUE)

# Path model
semPaths(
  fit_1,
  "std",
  layout = "tree2",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  cardinal = TRUE,
  edge.color = "black",
  residScale = 10
)

dev.copy(png,
         "./results/path_detearn1.png",
         width = 750,
         height = 600)
dev.off()

# Categorical model
fit_1c <- cfa(cfa_1,
              data = detearn,
              ordered = c("welljob",
                          "hardjob", "family", "child"))

summary(fit_1c, fit.measures = TRUE, standardized = TRUE)  #probs
```

# Full model

## Descriptives

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
names(merit_per_pref)
merit_per_pref %>% 
  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,race,gender,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

row.names(desc.issp) <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children"
)

save(desc.issp,file = "output/tables/desc.issp.Rdata")
```

## Measurement model

```{r}    
# 3. MEASURING BELIEFS & PERCEPTIONS ---------------------------------------------------------------
# 3.1 Define object --------------------------------------------------------------------------------
names(merit_per_pref)
# 3.2 Measurement continuous -----------------------------------------------------------------------

# Estimación del modelo simple
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- cfa(cfa_perpref1, data = merit_per_pref,
                    # missing = "ML",
                    ordered = c("wealthy","pareduc","ambition","hwork","people","polcone","race","gender",
                                "respons","yeduc","family","child","welljob","hardjob"))

names(merit_per_pref)
summary(fit_perpref1,
        fit.measures = TRUE,
        standardized = TRUE)
```

```{r}
plot_m_perpref1 <- '
	  # latent variables
	  percmerit =~ 1*hwork + ambition
	  percnmerit=~ 1*wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ 1*welljob + hardjob
	  prefnmerit=~ 1*family + child
	  '
plot_perpref1 <- cfa(plot_m_perpref1, data = merit_per_pref,
                    # missing = "ML"
                    )

nodeNames<- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Reasons for pay: well job",
  "Reasons for pay: hard job",
  "Reasons for pay: support family",
  "Reasons for pay: has children",
  "Perception \n meritocratic",
  "Perception \n non meritocratic",
  "Preference \n meritocratic",
  "Preference \n non meritocratic"
)

xfa <- 1*0.5

# Path model
semPaths(
  plot_perpref1 ,
  # whatLabels = "std", # Indica si aparecen los valores
  what = "mod",
  label.cex = 0.38, # Tamaño de las etiquetas dentro de los nodos.
  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
  residuals = T,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  nCharNodes = 0,
  curvePivot = F,
  curve = 3,
  rotation = 4,
  layout = "tree3",
  cardinal = "lat cov",
  legend.cex = 0.6, # 0.6
  label.cex = 4,
  label.font = 6,
  edge.label.font = 14,
  asize = 3,
  edge.width = 1.5,
  sizeMan = 6,# largo manifest
  sizeMan2= 3,# alto manifest
  sizeLat = 5, # largo de los circulos
  sizeLat2 = 11.5, # alto de los circulos
  residScale = 10,
  width =6, # 17
  height = 3, # 10
  # nodeNames = nodeNames,
  intercepts = F,
  reorder = T,
  thresholds = F,
  fixedStyle =1,
  node.height = 1,
  node.width = 3,
  label.scale = F,
  shapeMan = "rectangle",
  shapeLat = "ellipse",
  nodeLabels = nodeNames,
  details = T,
  mar = c(1,
          4, #left
          1, #top
          5), #right
  filetype ="pdf",
  filename = "output/images/meas-issp01"
  )->semplot01

```

```{r eval=FALSE, include=FALSE}
# Estimación del modelo c/ unmerit 2do orden
cfa_perpref2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit=~ parent + backg + network # 2nd order
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
fit_perpref2 <- cfa(cfa_perpref2, data = merit_per_pref,
                    missing = "ML")
summary(fit_perpref2,
        fit.measures = TRUE,
        standardized = TRUE)

# Path model
semPaths(
  fit_perpref2,
  layout = "tree",
  rotation = 2,
  intercepts = FALSE,
  style = "lisrel",
  curvePivot = TRUE,
  sizeMan = 3.5,
  sizeLat = 6
)

dev.copy(png,
         "./results/path_perpref2.png",
         width = 800,
         height = 650)
dev.off()

```

- Meritocratic perception:
  - How important is having ambition?
  - How important is hard work?

- Non-meritocratic perception
  - How important is coming from a wealthy family?
  - How important is having well-educated parents
  - How important is knowing the right people
  - How important is having political connections
  - How important is a person's race?
  - How important is being born a man or a woman?

- Meritocratic preferences
  - How well he or she does the job?
  - How hard he or she works at the job?

- Non-meritocratic preferences
  - What is needed to support a family?
  - Whether the person has children to support?

```{r}
labs <- c(
  "How important is having ambition?",
  "How important is hard work?",
  "How important is coming from a wealthy family?",
  "How important is having welleducated parents",
  "How important is knowing the right people",
  "How important is having political connections",
  "How important is a person's race?",
  "How important is being born a man or a woman?",
  "How well he or she does the job?",
  "How hard he or she works at the job?",
  "What is needed to support a family?",
  "Whether the person has children to support?"
)

tb.load<- data.frame(round(cbind(inspect(fit_perpref1,what="std")$lambda),digits = 2))
tb.load[tb.load==c(0.00)] <- NA

for (i in names(tb.load)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
}
tb.load[tb.load=="NA"] <- ""

#-------#
fm01<- data.frame(t(data.frame(fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"

#------chi2, df------#
fm04<- round(rbind(fm01),3)
fm04.1 <- fm04 %>% select(chisq,df) 
fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
fm04.1 <- select(fm04.1,"chi2df")
for (i in names(fm04.1)) {
  fm04.1[,i] <- as.character(fm04.1[,i])
}

#------CFI, RMSEA------#
fm04.2 <- fm04 %>% select(cfi,tli,rmsea) 
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

fm.df      <- bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(nobs(fit_perpref1)) 
fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")

#------ merge ------#
tb.fm<- bind_rows(tb.load,fm.df)
tb.fm<- tb.fm %>% mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$","$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>% select(vars,everything())
tb.perpref1 <- tb.fm
```

```{r}
save(tb.perpref1,file = "output/tables/tb.perpref1.RData")
```

# Other analysis

```{r}
# 3.3 Measurement categorical -pendiente- #####

# Estimación del modelo simple categórico
fit_1c <- cfa(
  cfa_1,
  data = merit_per_pref,
  ordered = c(
    "wealthy",
    "ambition",
    "hwork",
    "people",
    "polcone",
    "race",
    "gender",
    "pareduc",
    "welljob",
    "hardjob",
    "family",
    "child"
  )
)
summary(fit_1c, fit.measures = TRUE, standardized = TRUE)

# Estimación del modelo 2do orden categórico
fit_2c <- cfa(
  cfa_2,
  data = merit_per_pref,
  ordered = c(
    "wealthy",
    "ambition",
    "hwork",
    "people",
    "polcone",
    "race",
    "gender",
    "pareduc",
    "welljob",
    "hardjob",
    "family",
    "child"
  )
)
summary(fit_2c, fit.measures = TRUE, standardized = TRUE)
```

```{r eval=FALSE, include=FALSE}
# 3.4. Save factor scores --------------- #####

# Factor scores fit_perpref2

# Multiple imputation (para recuperar missings en scores) - REVIEW THIS, command lavPredict from 2017 should predict with missing values therefore saving from multiple imputation

# generate 5 multiple complete datasets
out <- mice(merit_per_pref[, c(3:10, 19:24)], m = 5)
D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(cfa_perpref2, data = D1)
fit2 <- cfa(cfa_perpref2, data = D2)
fit3 <- cfa(cfa_perpref2, data = D3)
fit4 <- cfa(cfa_perpref2, data = D4)
fit5 <- cfa(cfa_perpref2, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

# Generate factor scores
scores_dat = as.data.frame(scores)
dim(scores_dat)
str(scores_dat)
summary(scores)

# Generate object with factor scores
dim(merit_per_pref)
dim(scores_dat)
merit_per_pref_sco = cbind(merit_per_pref, scores)

# Save object with factor scores
names(merit_per_pref_sco)
merit_per_pref_sco = merit_per_pref_sco[, c(1:2, 25:31)]
save(merit_per_pref_sco,
     file = "./objects/merit_per_pref_sco.rda")

# Descriptives factor scores
load("./objects/merit_per_pref_sco.rda")
# Correlations
names(merit_per_pref_sco)
cormat = cor(merit_per_pref_sco[c(3, 7:9)],
             use = "pairwise.complete.obs")
corrplot.mixed(cormat)
#save
dev.copy(png,
         "./results/corplot_scores.png",
         width = 600,
         height = 600)
dev.off()

# Descriptive scatter countries
load("./objects/merit_per_pref_sco.rda")
agg_merit = merit_per_pref_sco %>% group_by(v5) %>% summarise_all(funs(mean))

agg_merit$v5c = countrycode(agg_merit$v5,
                            'iso3n', 'iso3c')

set_label(agg_merit$merit) = "Meritocratic perception"
set_label(agg_merit$bmerit) = "Meritocratic preferences"

sjp.scatter(
  agg_merit$merit,
  agg_merit$bmerit,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_mprefs.png",
         width = 600,
         height = 600)
dev.off()

sjp.scatter(
  agg_merit$merit,
  agg_merit$unmerit,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_munperc.png",
         width = 600,
         height = 600)
dev.off()

sjp.scatter(
  agg_merit$bmerit,
  agg_merit$need,
  dot.labels = agg_merit$v5c,
  fit.line = TRUE,
  show.ci = TRUE
)

dev.copy(png,
         "./results/mperc_munperc.png",
         width = 600,
         height = 600)
dev.off()
```

```{r}
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- cfa(cfa_perpref1, data = merit_per_pref,
                    # missing = "ML",
                    ordered = c("wealthy","pareduc","ambition","hwork","people","polcone","race","gender",
                                "respons","yeduc","family","child","welljob","hardjob"))
```

```{r eval=FALSE, include=FALSE}
### Check an alternative method
merit_per_pref$v5 <- as.factor(merit_per_pref$v5)
configural        <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5", estimator="ML")
weak.invariance   <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = "loadings", estimator="ML")
strong.invariance <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = c( "loadings", "intercepts"), estimator="ML")
strict.invariance <- cfa(cfa_perpref1, data=merit_per_pref, group = "v5",group.equal = c( "loadings", "intercepts", "residuals"), estimator="ML")
```

```{r eval=FALSE, include=FALSE}
anova(configural,weak.invariance,strong.invariance,strict.invariance)
```

```{r eval=FALSE, include=FALSE}
fitmeasures(configural       )
fitmeasures(weak.invariance  )
fitmeasures(strong.invariance) 
fitmeasures(strict.invariance) 
```

```{r eval=FALSE, include=FALSE}
# 3.5 Invariance full model ---pendiente- --------

fit.indices = c("chisq", "df", "cfi", "tli", "rmsea")

# Continuous first order
fit.conf1 = cfa(cfa_1,
                data = merit_per_pref,
                group = "v5",
                std.lv = TRUE)

fitMeasures(fit.conf1, fit.indices)
summary(fit.conf1)
inspect(fit.conf1, "theta")
inspect(fit.conf1, "cov.lv")

lavaan::
partialInvariance(fit_perpref1, type = "strict", p.adjust = "holm")
```