### 2021-06-02 processing/analysis_study1.Rmd

---
title: "Analysis Study 1"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
knitr::opts_knit$set(root.dir = "../")
options(scipen=9999) # desactivar notacion cientifica
rm(list=ls())
options(knitr.kable.NA = '')
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

## Libraries

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,sjPlot,lavaan,psych, knitr,kableExtra,ggplot2,stargazer,
               corrplot,semPlot)
```

```{r}
table_format = if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}

```

# Analysis

## Descriptives

- Meritocratic perception:
  - How important is having ambition?
  - How important is hard work?

- Non-meritocratic perception
  - How important is coming from a wealthy family?
  - How important is having well-educated parents
  - How important is knowing the right people
  - How important is having political connections
  - How important is a person's race?
  - How important is being born a man or a woman?

- Meritocratic preferences
  - How well he or she does the job?
  - How hard he or she works at the job?

- Non-meritocratic preferences
  - What is needed to support a family?
  - Whether the person has children to support?

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr,sjPlot,lavaan,psych, knitr,kableExtra,ggplot2,stargazer,
               corrplot,semPlot)
names(merit_per_pref)
merit_per_pref %>% 
  dplyr::select(ambition,hwork,wealthy,pareduc,people,polcone,race,gender,welljob,hardjob,family,child) %>% 
  psych::describe() %>% 
  as.data.frame() %>% 
  dplyr::select("Mean"=mean,"SD"=sd,"Min"=min,"Max"=max) %>% 
  round(2) ->desc.issp

row.names(desc.issp) <- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Important for pay: well job",
  "Important for pay: hard job",
  "Important for pay: support family",
  "Important for pay: has children"
)

# save(desc.issp,file = "output/tables/desc.issp.Rdata")
```

## Measurement model

```{r}    
# 3. MEASURING BELIEFS & PERCEPTIONS 
load(file = "input/data/proc/merit_per_pref.rda")
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
summary(fit_perpref1,
        fit.measures = TRUE,
        standardized = TRUE)
```

```{r, fig.width=10,fig.height=8}
load(file = "input/data/proc/merit_per_pref.rda")
plot_m_perpref1 <- '
	  # latent variables
	  percmerit =~ 1*hwork + ambition
	  percnmerit=~ 1*wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ 1*welljob + hardjob
	  prefnmerit=~ 1*family + child
	  '
plot_perpref1 <- lavaan::cfa(plot_m_perpref1, data = merit_per_pref)

nodeNames<- c(
  "Importance: hard work",
  "Importance: having ambition",
  "Importance: wealthy family",
  "Importance: educated parents",
  "Importance: race",
  "Importance: gender",
  "Importance: knowing people",
  "Importance: political connections",
  "Important for pay: well job",
  "Important for pay: hard job",
  "Important for pay: support family",
  "Important for pay: has children",
  "Perception \n meritocratic",
  "Perception \n non meritocratic",
  "Preference \n meritocratic",
  "Preference \n non meritocratic"
)

xfa <- 1*0.5

# Path model
semPlot::semPaths(
  plot_perpref1 ,
  # whatLabels = "std", # Indica si aparecen los valores
  what = "mod",
  label.cex = 0.38, # Tamaño de las etiquetas dentro de los nodos.
  edge.label.cex =  0.8, #  0.8 Tamaño de los valores estimados.
  residuals = T,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  nCharNodes = 0,
  curvePivot = F,
  curve = 3,
  rotation = 4,
  layout = "tree3",
  cardinal = "lat cov",
  legend.cex = 0.8, # 0.6
  label.cex = 4,
  label.font = 6,
  edge.label.font = 14,
  asize = 3,
  edge.width = 1.5,
  sizeMan = 6,# largo manifest
  sizeMan2= 3,# alto manifest
  sizeLat = 5, # largo de los circulos
  sizeLat2 = 11.5, # alto de los circulos
  residScale = 10,
  width =6, # 17
  height = 3, # 10
  # nodeNames = nodeNames,
  intercepts = F,
  reorder = T,
  thresholds = F,
  fixedStyle =1,
  node.height = 1,
  node.width = 3,
  label.scale = F,
  shapeMan = "rectangle",
  shapeLat = "ellipse",
  nodeLabels = nodeNames,
  details = T,
  mar = c(1,
          4, #left
          1, #top
          4), #right
  # filetype ="pdf",
  # filename = "output/images/meas-issp01"
  )->semplot01

```

```{r}
load(file = "input/data/proc/merit_per_pref.rda")
pacman::p_load(dplyr)
cfa_perpref1 <- '
	  # latent variables
	  percmerit =~ hwork + ambition
	  percnmerit=~ wealthy + pareduc +race + gender +people + polcone
	  prefmerit =~ welljob + hardjob
	  prefnmerit=~ family + child
	  '
fit_perpref1 <- lavaan::cfa(cfa_perpref1, data = merit_per_pref,
                            ordered = c("wealthy","pareduc","ambition","hwork",
                                        "people","polcone","race","gender",
                                        "respons","yeduc",
                                        "family","child",
                                        "welljob","hardjob"))
labs <- c(
  "How important is having ambition?",
  "How important is hard work?",
  "How important is coming from a wealthy family?",
  "How important is having welleducated parents",
  "How important is knowing the right people",
  "How important is having political connections",
  "How important is a person's race?",
  "How important is being born a man or a woman?",
  "How well he or she does the job?",
  "How hard he or she works at the job?",
  "What is needed to support a family?",
  "Whether the person has children to support?"
)

tb.load<- data.frame(round(cbind(lavaan::inspect(fit_perpref1,
                                                        what="std")$lambda),
                           digits = 2))
tb.load[tb.load==c(0.00)] <- NA

for (i in names(tb.load)) {
  # tb.load[,i] <- sjlabelled::as_character(tb.load[,i])
  tb.load[,i] <- sprintf(tb.load[,i], fmt = '%#.2f')
}
tb.load[tb.load=="NA"] <- ""
#-------#
fm01<- data.frame(t(data.frame(lavaan::fitmeasures(fit_perpref1, output ="matrix")[c("chisq","df","cfi","tli","rmsea"),]))); row.names(fm01) ="percnmerit"

#------chi2, df------#
fm04<- round(rbind(fm01),3)
fm04.1 <- fm04 %>% dplyr::select(chisq,df) 
fm04.1$chisq <- round(x = fm04.1$chisq,digits = 1)
fm04.1$df <- round(x = fm04.1$df,digits = 0)
fm04.1$chi2df <- paste0(fm04.1$chisq,"(",fm04.1$df,")")
fm04.1 <- dplyr::select(fm04.1,"chi2df")
for (i in names(fm04.1)) {
  fm04.1[,i] <- as.character(fm04.1[,i])
}

#------CFI, RMSEA------#
fm04.2 <- fm04 %>% dplyr::select(cfi,tli,rmsea) 
for (i in names(fm04.2)) {
  fm04.2[,i] <- sprintf(fm04.2[,i], fmt = '%#.3f')
}

fm.df      <- dplyr::bind_cols(fm04.1,fm04.2)
fm.df$nobs <- c(lavaan::nobs(fit_perpref1)) 
fm.df <- data.frame(t(fm.df)); colnames(fm.df) <- c("percnmerit")

#------ merge ------#
tb.fm<- dplyr::bind_rows(tb.load,fm.df)
tb.fm<- tb.fm %>% 
  dplyr::mutate(vars=c(labs,"$\\chi^2\\text{(df)}$","$\\text{CFI}$",
                       "$\\text{TLI}$","$\\text{RMSEA}$","$N$")) %>%
  dplyr::select(vars,everything())
tb.perpref1 <- tb.fm
```

```{r}
# save(tb.perpref1,file = "output/tables/tb.perpref1.RData")
```