### 2021-06-02 processing/preparation_study1.Rmd

---
title: "Data preparation: Study 1"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "show"
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
knitr::opts_knit$set(root.dir = "../")
options(scipen=9999) # desactivar notacion cientifica
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```

## Libraries
```{r}
if (!require("pacman")) install.packages("pacman") # install pacman
pacman::p_load(dplyr,lavaan,sjPlot,sjmisc,sjlabelled,haven,stringr,summarytools)
```

# Data

```{r}
issp2009 <- read_dta("../input/data/original/ZA5400_v4-0-0.dta")
```

```{r }
names(issp2009) <- tolower(names(issp2009))
# Check & generate IDs (v4: respondent, v5: country)
tapply(issp2009$v4, issp2009$v5, mean, na.rm = TRUE)
# Algunos paises (ej:826) tienen ids raros ...
# Generar nueva id secuencial individuos por país para merges
issp2009$id <- sequence(tabulate(issp2009$v5))
tapply(issp2009$id, issp2009$v5, mean, na.rm = TRUE) # ok
summary(issp2009$id)
issp2009$idtem = issp2009$v5 * 10000
issp2009$id = issp2009$idtem + issp2009$id
```

## Perceptions (getting-ahead)

*  v6: wealthy family
*  v7: educated parents
*  v9: ambition 
* v10: hard work 
* v11: right people
* v12: political connections
* v14: race 
* v16: gender 
* IDs: v5 (country) & id (individual id)

```{r}
# names(issp2009)          #Check names
get_label(issp2009[8:18])#Check variable labels

# 1. Select and generate new variables
getahead <- 
  issp2009 %>%
  select("id","v5","v6","v7","v9","v10","v11","v12","v14","v16") %>% 
  mutate(wealthy = v6,
         pareduc = v7,
         ambition = v9,
         hwork = v10,
         people = v11,
         polcone = v12,
         race = v14,
         gender = v16) %>% 
  select(-c("v6","v7","v9","v10","v11","v12","v14","v16"))

names(getahead)

#2. Recode missing values
# * 8 & 9 to NA
frq(getahead$wealthy)
getahead[3:10][getahead[3:10] > 5] <- NA
frq(getahead$gender) # OK
summary(getahead)    # OK

#3. value labels
get_labels(getahead$wealthy)
getahead[, 3:10] <- 
  set_na(x = getahead[, 3:10],
         na = c(8,9))
get_labels(getahead[3:10]) # OK
frq(getahead[3:10])        # OK

# 4. variable label

getahed.label <-  str_remove_all(string = get_label(getahead[3:10]),
                                 pattern = "Q1[abcdefghijk] ")
getahed.label

getahead[3:10] = 6 - getahead[3:10] # Reverse code for SEM models
getahead[3:10] <- set_label(getahead[3:10],label = getahed.label) # variable label
get_label(getahead[3:10]) # OK

# 5. final ckeck   
view_df(getahead,show.frq = T, show.prc = T)
```

## Preferences (detearn)

```{r}	    
# names(issp2009)

# 1. Select and generate new variables
detearn <- 
  issp2009 %>%
  select("id","v5","v47", "v48", "v49", "v50", "v51","v52") %>% 
  mutate(respons 	= v47,
         yeduc  	= v48,
         family   = v49,
         child    = v50,
         welljob  = v51,
         hardjob  = v52) %>% 
  select(-c("v47", "v48", "v49", "v50", "v51","v52"))

#2. Recode missing values
# * 8 & 9 to NA
summary(detearn)
frq(detearn$respons)
detearn[3:8][detearn[3:8] > 5] <- NA
frq(detearn$respons)

# Out Portugal (no answer in well job item)
detearn[which(detearn$v5 == 620), ] = NA

#3. value labels
get_labels(detearn[3:8])
detearn[, 3:8] <- 
  set_na(detearn[, 3:8],
         na =  c(8, 9)) 
summary(detearn)      # Value 0 in Portugal Answers
frq(detearn$welljob)  # Check

detearn$welljob[detearn$welljob == 0] = NA # recode to NA value 0 in welljob
get_labels(detearn$welljob)
frq(detearn$welljob)

# value label correction in welljob
detearn$welljob <- set_labels(x = detearn$welljob,
                              labels = get_labels(detearn$hardjob))
frq(detearn$welljob) # OK, solved
frq(detearn[3:8])    # Check value labels

# 4. Variable label
get_label(detearn[3:8])
detearn.label <-  str_remove_all(string = get_label(detearn[3:8]),
                                 pattern = "Q12[abcdef] ")

detearn[3:8] <-  6 - detearn[3:8] # Reverse code for SEM models
detearn[3:8] <- set_label(detearn[3:8],label = detearn.label) # variable label

frq(detearn[3:8])

# 5. final ckeck   
view_df(detearn,show.frq = T, show.prc = T)
```

# Merge

```{r}
merit_per_pref   <- left_join(getahead, detearn, by = c("id", "v5"))

```

# Summary
```{r,results='asis', echo=FALSE}
st_css() # esto configura el css de dfsummary
print(dfSummary(merit_per_pref),
      method = "render")
```

```{r}
save(merit_per_pref, file = "../input/data/proc/merit_per_pref.rda")
```