### 2021-06-03 processing/preparation_study3.rmd

---
title: "Data preparation Study 3 - Merit Scale"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
# Knitr options

knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

This is the preparation code of the data for Study 3:  Additional validity analyses. The prepared data is `data_s3_conv.RData` in "Convergence section" and `data_s3_conv.RData` in "Invariance section"

# Convergence

## Libraries
```{r}
pacman::p_load(haven, sjPlot, dplyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
```

## Load original data

```{r}
dat04w3 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
dim(data) 
```

## Create Perceptions and Preferences for Meritocracy (PPM-S) factor scores

The following code creates new variables corresponding to the four dimensions of the PPM scale. Based on the measurement model, four factorial scores are created. To retrieve cases the "mice" function is used

```{r results='hide'}
dat04w3

dat04w3$meritv03_perc_effort <- as.numeric(dat04w3$meritv03_perc_effort)
dat04w3$meritv03_perc_talent <- as.numeric(dat04w3$meritv03_perc_talent)
dat04w3$meritv03_perc_netw <- as.numeric(dat04w3$meritv03_perc_netw)
dat04w3$meritv03_perc_wpart <- as.numeric(dat04w3$meritv03_perc_wpart)
dat04w3$meritv03_pref_effort <- as.numeric(dat04w3$meritv03_pref_effort)
dat04w3$meritv03_pref_talent <- as.numeric(dat04w3$meritv03_pref_talent)
dat04w3$meritv03_pref_wpart <- as.numeric(dat04w3$meritv03_pref_wpart)
dat04w3$meritv03_pref_netw <- as.numeric(dat04w3$meritv03_pref_netw)

model04 <- '
perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw'

# generate 5 multiple complete datasets
out <- mice(dat04w3[23:30], m = 5)

D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(model04, data = D1)
fit2 <- cfa(model04, data = D2)
fit3 <- cfa(model04, data = D3)
fit4 <- cfa(model04, data = D4)
fit5 <- cfa(model04, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

#Generate factor scores
scores_dat = as.data.frame(scores)

# Adaptar base para agregar factor scores (sacar todos NA)
dat04w3 = dat04w3[rowSums(is.na(dat04w3[23:28])) != 5, ]

# Merge with factor scores
dat04w3 = cbind(dat04w3, scores)

# Check
stargazer(dat04w3[, 78:81], type = "text")

```

## Create The Neoliberal Beliefs Inventory (NBI) factor scores

The following code creates new variables corresponding to the four dimensions of the NBI scale. Based on the measurement model, four factorial scores are created. To retrieve cases the "mice" function is used
```{r results='hide'}
datnbmerit <- dat04w3 %>% select(nb_merit_1,
nb_merit_2,
nb_merit_3,
nb_merit_4,
nb_merit_5,
nb_merit_6,
nb_merit_7,
nb_merit_8)
psych::alpha(datnbmerit)

dat04w3$nb_merit_1 <- as.numeric(dat04w3$nb_merit_1)
dat04w3$nb_merit_2 <- as.numeric(dat04w3$nb_merit_2)
dat04w3$nb_merit_3 <- as.numeric(dat04w3$nb_merit_3)
dat04w3$nb_merit_4 <- as.numeric(dat04w3$nb_merit_4)
dat04w3$nb_merit_5 <- as.numeric(dat04w3$nb_merit_5)
dat04w3$nb_merit_6 <- as.numeric(dat04w3$nb_merit_6)
dat04w3$nb_merit_7 <- as.numeric(dat04w3$nb_merit_7)
dat04w3$nb_merit_8 <- as.numeric(dat04w3$nb_merit_8)

model04 <- '
nbmerit =~  nb_merit_1+
nb_merit_2+
nb_merit_3+
nb_merit_4+
nb_merit_5+
nb_merit_6+
nb_merit_7'

# generate 5 multiple complete datasets
out <- mice(dat04w3[57:64], m = 5)

D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(model04, data = D1)
fit2 <- cfa(model04, data = D2)
fit3 <- cfa(model04, data = D3)
fit4 <- cfa(model04, data = D4)
fit5 <- cfa(model04, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

#Generate factor scores
scores_dat = as.data.frame(scores)

# Adaptar base para agregar factor scores (sacar todos NA)
dat04w3 = dat04w3[rowSums(is.na(dat04w3[57:64])) != 5, ]

# Merge with factor scores
dat04w3 = cbind(dat04w3, scores)

data_s3_conv <- dat04w3 %>% select(nbmerit, perc_merit,pref_merit,perc_nmerit,pref_nmerit,get_ah_1, get_ah_5) %>% na.omit()

```

## Labels

### Variable labels

Created variables are labeled and described

```{r}

set_label(data_s3_conv$nbmerit)  <- "Personal wherewithal - NBI"
set_label(data_s3_conv$perc_merit)  <- "Meritocratic perception"
set_label(data_s3_conv$pref_merit)   <- "Meritocratic preferences"
set_label(data_s3_conv$perc_wpart)    <- "Unmeritocratic perception"
set_label(data_s3_conv$pref_nmerit)  <-  "Unmeritocratic preferences"

set_label(data_s3_conv$get_ah_1)  <- "Importance of a rich family to get ahead"
set_label(data_s3_conv$get_ah_5)   <- "Importance of hard work to get ahead"

sjmisc::descr(data_s3_conv)

```

## Final summary
```{r}
sjmisc::descr(data_s3_conv,
      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
      kable(.,"markdown")

print(dfSummary(data_s3_conv, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")

```

## save data

```{r}

save(data_s3_conv,file = "input/data/proc/data_s3_conv.RData")
```

# Invariance

## Libraries
```{r}
pacman::p_load(haven, sjPlot, dplyr, lavaan, stargazer, corrplot, knitr, kableExtra, summarytools, sjlabelled,Hmisc,psych,tidyr)

```

## Load longitudinal data

```{r}
w01 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola1.sav",verbose = F)
w02 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola2.sav",verbose = F)
w03 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
```

## Preparation wave 1

Add tags wave 1
```{r }
label(w01$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
w01$Intro <- set_labels(w01$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 10}
w01_Intro <- w01 %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w01_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w01b <- w01 %>% filter(Intro==1, Finished==1) %>% as.data.frame()
w01b$ID<- stringr::str_split_fixed(w01b$ticket,"_", 4)[,1] # En la variable ticket el primer código es el ID, todo lo demás se borra
```

Frequency of respondents who Accept and Finish wave 1.
```{r fig.width= 20 }
w01b$Intro <- set_labels(w01b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w01_Introb <- w01b %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w01_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

A variable is created to indicate the order in which the items on the PPM scale are presented. The variables are renamed and those of interest are selected. The results are described.
```{r}
w01c <- w01b
w01c$grupo <- NA
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c[w01c==0] <- NA
w01c <- w01c %>% select(grupo,
                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w01c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation wave 2

Add tags wave 2
```{r }
label(w02$Intro) <- "If you wish to participate, select the alternative I ACCEPT take part in this investigation. If you do not want to participate in this study, press the alternative I NOT ACCEPT take part in this investigation."
w02$Intro <- set_labels(w02$Intro,  labels = c("I ACCEPT take part in this investigation","I NOT ACCEPT take part in this investigation."))
```

Frequency of respondents who Accept and Finish wave 2.
```{r}
w02_Intro <- w02 %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w02_Intro, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

Delete those who do not accept and Delete those who do not accept
```{r}
w02b <- w02 %>% filter(Intro==1, Finished==1) %>% as.data.frame()
w02b$ID <- stringr::str_split_fixed(w02b$ticket, "_", 4)[,1]
```

Frequency of respondents who Accept and Finish wave 2.
```{r }
w02b$Intro <- set_labels(w02b$Intro,  labels = c("I ACCEPT take part in this investigation", "I NOT ACCEPT take part in this investigation."))
w02_Introb <- w02b %>% select(Intro,Finished) %>% as.data.frame()
sjPlot::view_df(w02_Introb, show.type = T,show.frq = T, show.string.values = T, verbose = F, show.id = F)
```

A variable is created to indicate the treatment to which the respondent was exposed. The variables are renamed and those of interest are selected. The results are described.

```{r}
w02c <- w02b
w02c$treat <- NA
w02c$treat[w02c$FL_6_DO_Control==1] <- 1
w02c$treat[w02c$FL_6_DO_Tratamiento==1] <- 2
w02c$treat[w02c$FL_6_DO_Tratamiento_desigualdad==1] <- 3
w02c$treat <- factor(w02c$treat,levels = c(1,2,3),labels = c("control","pobreza","desiguald"))
w02c <- w02c %>% rename(perc_effort_w2=meritv03_perc_effort,
                        perc_talent_w2=meritv03_perc_talent,
                        perc_wpart_w2 =meritv03_perc_wpart,
                        perc_netw_w2  =meritv03_perc_netw,
                        pref_effort_w2=meritv03_pref_effort,
                        pref_talent_w2=meritv03_pref_talent,
                        pref_wpart_w2 =meritv03_pref_wpart,
                        pref_netw_w2  =meritv03_pref_netw)
w02c <- w02c %>% select(perc_effort_w2,perc_talent_w2,perc_wpart_w2,perc_netw_w2,
                        pref_effort_w2,pref_talent_w2,pref_wpart_w2,pref_netw_w2,
                        sexo_w2=sexo,edad_w2=edad,edcep_w2=edcep, treat, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w02c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Preparation wave 3

```{r}
table(w03$Finished==1)
w03b <- filter(w03,Finished==1) # Elimina los que no aceptan
dim(w03b)
w03b$ID <- stringr::str_split_fixed(w03b$ticket, "_", 4)[,1]
```

The variables of the PPM scale are renamed indicating the wave. Those of interest are selected. The results are described.
```{r}
w03c <- w03b
w03c <- w03c %>% rename(perc_effort_w3=meritv03_perc_effort,
                        perc_talent_w3=meritv03_perc_talent,
                        perc_wpart_w3 =meritv03_perc_wpart,
                        perc_netw_w3  =meritv03_perc_netw,
                        pref_effort_w3=meritv03_pref_effort,
                        pref_talent_w3=meritv03_pref_talent,
                        pref_wpart_w3 =meritv03_pref_wpart,
                        pref_netw_w3  =meritv03_pref_netw)
w03c <- w03c %>% select(perc_effort_w3,perc_talent_w3,perc_wpart_w3,perc_netw_w3,
                        pref_effort_w3,pref_talent_w3,pref_wpart_w3,pref_netw_w3,
                        sexo_w3=sexo,edad_w3=edad,edcep_w3=edcep, ID)
#-----------------------------------------------------------------------#
print(dfSummary(w03c, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

## Merge wide data

```{r}
wide01 <- inner_join(x = w01c,y = w02c, by = "ID") 
dim(wide01) #en w01 y w02
```

```{r}
wide02 <- inner_join(x = wide01,y = w03c, by = "ID") 
dim(wide02) #en w01, w02 y w03
```

```{r}
print(dfSummary(w02, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")
```

```{r}
#save(wide01,file = "input/data/proc/datawide.RData")
#save(wide02,file = "input/data/proc/datawide_123.RData")
```

## Wide to long data

```{r}
library(datasets)
library(data.table)
wide <- setDT(wide02)
long01<- melt(wide,variable.name = "wave",
              measure = patterns("^perc_effort_","^perc_talent_","^perc_wpart_","^perc_netw_",
                                       "^pref_effort_","^pref_talent_","^pref_wpart_","^pref_netw_",
                                       "^sexo_","^edad_","^edcep_"),
              value.name = c("perc_effort","perc_talent","perc_wpart","perc_netw",
                             "pref_effort","pref_talent","pref_wpart","pref_netw",
                             "sexo","edad","edcep"))
#save(long01,file = "input/data/proc/datalong.RData")

```

## Factorial survey data:

Prepare poverty experiment data (wave 01):

```{r}
w01c <- w01 %>% filter(Intro==1) %>% as.data.frame()
w01c$ID<- stringr::str_split_fixed(w01c$ticket,"_", 4)[,1]
w01c$grupo <- NA
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v01==1] <- 1
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v02==1] <- 2
w01c$grupo[w01c$FL_21_DO_merit_perc_pref_julio19v03==1] <- 3
w01c$perc_effort_w1 <- rowSums(w01c[,c(matches(match = "perc_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_talent_w1 <- rowSums(w01c[,c(matches(match = "perc_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$perc_wpart_w1  <- rowSums(w01c[,c(matches(match = "perc_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$perc_netw_w1   <- rowSums(w01c[,c(matches(match = "perc_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_effort_w1 <- rowSums(w01c[,c(matches(match = "pref_effort",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_talent_w1 <- rowSums(w01c[,c(matches(match = "pref_talent",vars = names(w01c)))],na.rm = TRUE)
w01c$pref_wpart_w1  <- rowSums(w01c[,c(matches(match = "pref_wpart" ,vars = names(w01c)))],na.rm = TRUE)
w01c$pref_netw_w1   <- rowSums(w01c[,c(matches(match = "pref_netw"  ,vars = names(w01c)))],na.rm = TRUE)
w01c[w01c==0] <- NA
w01c <- w01c %>% select(grupo,
                        perc_effort_w1,perc_talent_w1,perc_wpart_w1,perc_netw_w1,
                        pref_effort_w1,pref_talent_w1,pref_wpart_w1,pref_netw_w1,
                        sexo_w1=sexo,edad_w1=edad,edcep_w1=edcep, ID)
```

```{r}
w01fs <- w01c
names(w01fs)<- stringr::str_replace_all(string = names(w01fs),pattern = "_w1",replacement = "") 
```

Prepare factorial survey data (wave 02):

```{r}
# fs01 <- sjlabelled::read_spss("input/data/original/factorial_ola2.sav",verbose = F)
load(file = "papelera/input/data/original/factorial_ola2.RData") 
```

```{r}
names(long01)
fs02 <- fs01 %>% 
  filter(consen==1) %>% 
  mutate(grupo=4,ID=ResponseId,) %>% 
  select(grupo,ID,"perc_effort"="merit_perc_effort",
         "perc_talent" = "merit_perc_talent",
         "perc_wpart" = "merit_perc_wpart",
         "perc_netw"  = "merit_perc_netw",
         "pref_effort" = "merit_pref_effort",
         "pref_talent" = "merit_pref_talent",
         "pref_wpart" = "merit_pref_wpart",
         "pref_netw"  = "merit_pref_netw",
         sexo,edad,
         edcep=educ)
```

Identify dataset of origin:

```{r}
w01fs$dataset <- "pvw01"
fs02$dataset  <- "fsw02"
```

Merge data sets:

```{r}
data_s3_inv<-bind_rows(fs02,w01fs) %>%
  set_label(label = get_label(fs02)) %>%
  set_labels(labels = get_labels(fs02))
```

```{r}
save(data_s3_inv,file = "input/data/proc/data_s3_inv.RData")
```