### 2021-08-11 index.Rmd
unir estudios & pendientes

---
title: "PPM-S"
csl: "input/bib/apa6.csl"
bibliography: "input/bib/meritocracy.bib"
author: " "
output:
    bookdown::html_document2:
            smart: false
            number_sections: false
---

Last update: `r format(Sys.time(), "%A %d %B %Y %X")`

```{r, child="readme.md", eval=TRUE}

```