### 2022-12-26 papelera/previous-submission/sjr/Title Page.Rmd
modificaciones paper-blinded para nuevo envío a SIR - primera parte hasta methods

---
title: "Measuring Perceptions and Preferences for Meritocracy"
linestretch: '1.5'
output:
  bookdown::pdf_document2:
     template: null
     toc: false
     number_sections: false
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 10pt
---

Juan Carlos Castillo, Department of Sociology, Universidad de Chile (
juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile )

Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile

Jorge Atria, Department of Sociology, Universidad Diego Portales

Julio Iturra, Center for Social Conflict and Cohesion Studies

Francisco Meneses, Department of Sociology, Universidad de Chile