### 2023-02-02 Title Page.Rmd
ajustes submit

---
title: "Measuring Perceptions and Preferences for Meritocracy [^*]"
linestretch: '1.5'
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    number_sections: false
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 12pt

---

Juan Carlos Castillo, Department of Sociology, Universidad de Chile *

Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile

Jorge Atria, Department of Sociology, Universidad Diego Portales

Julio Iturra, Bremen International Graduate School of Social Sciences

Francisco Meneses, Department of Sociology, Universidad de Chile

Word count: 8998

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors recognize the support of ANID/FONDECYT grant 1160921 and of  the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.