### 2023-05-16 Title Page.Rmd

---
title: "Measuring Perceptions and Preferences for Meritocracy [^*]"
linestretch: '1.5'
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    number_sections: false
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 12pt

---

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile. *

ORCID: 0000-0003-1265-7854

Julio Iturra, Bremen International Graduate School of Social Sciences, Bremen, Germany.

ORCID: 0000-0003-1993-3479

Luis Maldonado, Institute of Sociology, Pontificia Universidad Católica de Chile, Santiago de Chile.

ORCID: 0000-0002-0028-4766

Jorge Atria, Department of Sociology, Universidad Diego Portales, Santiago de Chile.

ORCID: 0000-0001-7145-3048

Francisco Meneses, Department of Sociology, Universidad de Chile, Santiago de Chile.

Word count: 8998

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) -  Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile). The authors recognize the support of ANID/FONDECYT grant 1210847 and of the Center for Conflict and Social Cohesion Studies (COES)  grant ANID/FONDAP 1513009.