### 2023-05-18 processing/preparation_study1_2019.Rmd
submission IJS

---
title: "Data preparation: Study 1_2019"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "show"
editor_options:
  chunk_output_type: console
---

## Note: this is an update  of the original study_1 (which used 2009 data) Here we use the last available file from wave 2019.

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
knitr::opts_knit$set(root.dir = "../")
options(scipen=9999) # desactivar notacion cientifica
```

```{css, echo=FALSE}
# /* expandir el espacio del documento*/
body .main-container {
      max-width: 1600px !important;
      width: 1600px !important;
    }
#/* expandir width de la TOC*/
div.tocify {
    width: 22% !important;
    max-width: 331px !important;
    max-height: 102% !important;
}
```
This is the preparation code of the data for Study 1_2019: Measuring meritocracy with international secondary data. The prepared data is `data_s1_2019.RData`

## Libraries
```{r}
if (!require("pacman")) install.packages("pacman") # install pacman
pacman::p_load(dplyr,lavaan,sjPlot,sjmisc,sjlabelled,haven,stringr,summarytools)
```

# Data

```{r}
issp2019 <- read_dta("input/data/original/ZA7600_v3-0-0.dta")
```

First, we check country identification variables and generate a specific individual-country variable.

```{r }
names(issp2019) <- tolower(names(issp2019))
find_var(data = issp2019,"country") # country and c_sample
find_var(data = issp2019,"id") #caseid
dplyr::n_distinct(issp2019$caseid) # ok, id única por caso

# Check & generate IDs (caseid: respondent, country: country)
tapply(issp2019$country, issp2019$caseid, mean, na.rm = TRUE)
# Algunos paises (ej:826) tienen ids raros ...
# Generar nueva id secuencial individuos por país para merges
issp2019$id <- sequence(tabulate(issp2019$country))
tapply(issp2019$id, issp2019$country, mean, na.rm = TRUE) # ok
summary(issp2019$id)
issp2019$idtem = issp2019$country * 10000
issp2019$id = issp2019$idtem + issp2019$id
```

## Perceptions (getting-ahead)

The following variables are those who are conceptualized as meritocratic and Non-meritocratic perception.

*  v1: wealthy family
*  v2: educated parents
*  XX: ambition 
*  v4: hard work 
*  v5: right people
*  v6: political connections
*  v8: race 
* v10: gender 
* IDs: v5 (country) & id (individual id)

First, we rename the variables to a more substantial name. In second place, we check for missing values and declare them as NA. Then, we reverse code all the variables, because number 1 represents the "Essential" and 5 represents "Not important at all". Finally, whe adjust variable labels to a more intuitive one.

```{r}
# names(issp2009)          #Check names
get_label(issp2009[8:18])#Check variable labels

# 1. Select and generate new variables
getahead <- 
  issp2009 %>%
  select("id","v5","v6","v7","v9","v10","v11","v12","v14","v16") %>% 
  mutate(wealthy = v6,
         pareduc = v7,
         ambition = v9,
         hwork = v10,
         people = v11,
         polcone = v12,
         race = v14,
         gender = v16) %>% 
  select(-c("v6","v7","v9","v10","v11","v12","v14","v16"))

names(getahead)

#2. Recode missing values
# * 8 & 9 to NA
frq(getahead$wealthy)
getahead[3:10][getahead[3:10] > 5] <- NA
frq(getahead$gender) # OK
summary(getahead)    # OK

#3. value labels
get_labels(getahead$wealthy)
getahead[, 3:10] <- 
  set_na(x = getahead[, 3:10],
         na = c(8,9))
get_labels(getahead[3:10]) # OK
frq(getahead[3:10])        # OK

# 4. variable label

getahed.label <-  str_remove_all(string = get_label(getahead[3:10]),
                                 pattern = "Q1[abcdefghijk] ")
getahed.label

getahead[3:10] = 6 - getahead[3:10] # Reverse code for SEM models
getahead[3:10] <- set_label(getahead[3:10],label = getahed.label) # variable label
get_label(getahead[3:10]) # OK

# 5. final ckeck   
view_df(getahead,show.frq = T, show.prc = T)
```

## Preferences (detearn)

* v47: responsibility job
* v48: year on education
* v49: support family
* v50: has children to support
* v51: well job
* v52: hard job
* IDs: v5 (country) & id (individual id)

Same as the meritocratic perception. First, we rename the variables to a more substantial name. In second place, we check for missing values and declare them as NA. Then, we reversecode al the variables, because number 1 represents the "Esential" and 5 represents "Not important at all". Finally, whe adjust variable labels to a more intuitive one.

```{r}	    
# names(issp2009)

# 1. Select and generate new variables
detearn <- 
  issp2009 %>%
  select("id","v5","v47", "v48", "v49", "v50", "v51","v52") %>% 
  mutate(respons 	= v47,
         yeduc  	= v48,
         family   = v49,
         child    = v50,
         welljob  = v51,
         hardjob  = v52) %>% 
  select(-c("v47", "v48", "v49", "v50", "v51","v52"))

#2. Recode missing values
# * 8 & 9 to NA
summary(detearn)
frq(detearn$respons)
detearn[3:8][detearn[3:8] > 5] <- NA
frq(detearn$respons)

# Out Portugal (no answer in well job item)
detearn[which(detearn$v5 == 620), ] = NA

#3. value labels
get_labels(detearn[3:8])
detearn[, 3:8] <- 
  set_na(detearn[, 3:8],
         na =  c(8, 9)) 
summary(detearn)      # Value 0 in Portugal Answers
frq(detearn$welljob)  # Check

detearn$welljob[detearn$welljob == 0] = NA # recode to NA value 0 in welljob
get_labels(detearn$welljob)
frq(detearn$welljob)

# value label correction in welljob
detearn$welljob <- set_labels(x = detearn$welljob,
                              labels = get_labels(detearn$hardjob))
frq(detearn$welljob) # OK, solved
frq(detearn[3:8])    # Check value labels

# 4. Variable label
get_label(detearn[3:8])
detearn.label <-  str_remove_all(string = get_label(detearn[3:8]),
                                 pattern = "Q12[abcdef] ")

detearn[3:8] <-  6 - detearn[3:8] # Reverse code for SEM models
detearn[3:8] <- set_label(detearn[3:8],label = detearn.label) # variable label

frq(detearn[3:8])

# 5. final ckeck   
view_df(detearn,show.frq = T, show.prc = T)
```

# Merge

We merge the meritocratic perceptions and preferences variables into a single dataset.

```{r}
data_s1   <- left_join(getahead, detearn, by = c("id", "v5"))
```

# Final summary
```{r,results='asis', echo=FALSE}
st_css() # esto configura el css de dfsummary
print(dfSummary(data_s1),
      method = "render")
```

```{r}
save(data_s1, file = "input/data/proc/data_s1.RData")
```